\pagenumbering{arabic}

\nomenclature{$\delta_{rq}$}{Binary variable attributing a quality to a resource}
\nomenclature{$k_{prq$}}{Coefficient of resource production}
\nomenclature{$g$}{Acceleration due to gravity}

# Introduction

```{r Setup, include=FALSE, results="hide", warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "figures/", dev=c("cairo_pdf"), fig.pos='htbp', fig.width=5, fig.height=3.5, dpi=700, fig.show='hold', fig.lp="fig:", cache=FALSE, par=TRUE, echo=FALSE, results="hide", message=FALSE, warning=FALSE)

library(extrafont)
font_import(pattern="LinBiolinum")

library(ggplot2)
```

\epigraph{``Cities are the crucible of human civilization, the drivers towards potential disaster, and the source of the solution to humanity's problems.''}{\citet{Bettencourt2010}, \emph{Nature}}

\epigraph{``\textbf{system}, n. An organized or connected group of things \ldots A group or set of related or associated things perceived or thought of as a unity or complex whole.''}{\citet{OED-system2016}}

Cities have brought a high quality to billions, by enabling the efficient provision of goods and services. Therefore, in many ways, it is a cause for celebration that the global urban population is growing, as more and more people find in them access to greater wealth, improved healthcare, and other socioeconomic services [@WorldBank2009]. However, to support their populations, cities consume and produce unsustainable quantities energy, water, and waste. As such, efforts are being devoted to reducing urban resource consumption and waste output.

<!--
Not only is the world's urban population growing, but so is our understanding of the nature of cities. Whilst being complicated amalgamations of people, businesses, institutions and geographical features, cities are surprisingly comprehensible
-->

Cities face many challenges, and thus questions about how they should be designed, developed, governed, and organised, are increasingly important. This thesis considers these questions in relation to concerns about levels of consumption and waste in cities. In addressing these challenges, there are many points at which planners and policy makers might make decisions in the vast networks of infrastructure and chains of processes which cities use to convert and supply goods and services.

Consider a city's electricity supply, in which a power plant burns a fossil-fuel and uses the resulting heat energy to generate steam which turns a turbine to generate electrical energy which is then delivered to the consumer by cable (Figure \ref{fig:elec}). Coal could be saved by encouraging consumers to use less electricity. Another approach could be to increase power plant efficiencies, so as to minimise energy losses within the plant. Alternatively, the power plant could be replaced by a process using renewable resources, such as solar. While interventions such as these have their place, this thesis will zoom out to look at how resource management works as a whole. Is the overall mix of power plants, water treatment facilities, waste management procedures, and other activities, the most efficient way to deliver resources?

\begin{figure}
	\centering
	\begin{spacing}{1.0}
	\sffamily
	\resizebox{\columnwidth}{!}{\input{01_introduction/power-network.pdf_tex}}
	\end{spacing}
	\caption{Some components of urban electricity system.}
	\label{fig:elec}
\end{figure}

In meeting consumer demand for electricity, the coal, power plant, electricity, and power lines are but a few components in the network of infrastructure used to manage a city's electricity. This network could include other energy sources (renewable and non-renewable), at various scales (large centralised power plants at one extreme, and small distributed microgeneration units at the other), all of which work in combination to meet electricity demand at any point in time. In other words, a city's electricity supply is a good example of a 'system' as defined above. Furthermore, the electricity system is but one part of the larger system which manages a city's energy, and this system is in turn just one of several systems at work in a city, others including water management, waste management, and transport. This thesis considers how system improvements can benefit urban resource management. More specifically, why and how should systems for the energy, water and waste sectors -- traditionally considered separately -- work together?

<!--
seen for example in campaigns to encourage behavioural change in consumers (e.g. turn their thermostats down), and policy measures to introducing new technologies (e.g. renewable energy). Another approach is to better design and operate the 'systems' which manage the resources. The management of urban resources requires networks of infrastructure to convert and transport resources. 

While the resource quantities involved in cities are large, they manage their resources more efficiently than smaller types of settlement. This is because the infrastructure in cities can be 
Thus while [TODO: stat], 
and thus, an urban dweller will typically use resources more efficiently than a rural dweller who has the same quality of life.

Today, over [TODO:] per cent of us live in urban areas (towns or cities), and this proportion is expected to rise to over [TODO:] per cent by [TODO:] [TODO: citation].

One particular reason for the success of cities is in how they can manage the resources on which we rely.  

resources to their point of need more efficiently than smaller types of settlement. [POSSIBLY: The same mechanisms by which firms can organise enable ... ]


- But presents challenges for resource consumption and generation
- But also opportunities
- The research field which studies the ins and outs of urban resource management is UM, which is [brief summary]

but there is one opportunity specific to cities which merits further study, namely the 

The rapid urbanisation of the globe is presenting both challenges and opportunities in the way that the Earth's finite resources are used in view of the problems presented by resource depletion and climate change. This project seeks firstly to take a broad look at global urbanisation trends and the environmental and economic challenges presented with respect to resource management; before looking at opportunities in the urban environment for more efficient use of resources through resource integration. Urban metabolism is introduced as the underlying theoretical concept which can be used to assess how efficiently an area uses resources. Having considered this big picture, the project focuses specifically on the development of mathematical models which optimise the integrated provision and management of energy, water and waste resources in an urban area. Existing models in these fields are reviewed, before the first steps in developing an integrated resource model are described. It is anticipated that the final outcome of the study will be a suite of tools that work at various scales, with different degrees of temporal and spatial resolution. This will be useful to urban planners, policy makers and utility service companies. First, global trends in urbanisation are considered.

TODO: complete Introduction, e.g. with how our development of UM theory fits in.
-->

## Urbanisation: the challenge

Whatever mix of approaches is adopted to better manage the world's resources -- whether they be behavioural, efficiency, technological, or system wide -- the need to better manage urban resources is evident by the following argument:

1. The world is urbanising
2. Urbanisation ...
	a. ... is on the whole a desirable phenomenon,
	b. ... leads to environmentally and economically unsustainable resource consumption and output
3. Therefore, urbanisation should not be opposed, but we should nevertheless act to reduce urban resource consumption and waste.

The rest of this Section considers the evidence in support of this argument.

### Urbanisation: an unstoppable and desirable phenomenon

Currently, urban areas are home to just over half of the global population [@UN2014]. By 2030, it is expected that the urban population grow to 60 per cent of the total, and to over 65 per cent by 2050 (Figure \ref{fig:pop_forecasts}).}. This growth has three components: cities absorb population growth; migration from rural areas; rural areas become reclassified as urban [@Cohen2006]. Growth predictions are not just the view of a small number, but widely accepted opinion. Thus even @Cohen2004 who takes a skeptical view of the projections (pointing out that historical forecasts can be overestimates by more than 75 per cent, and that there is no universally consistent definition of 'urban'[^urban-def]) concludes:

> *"...despite all the problems of error and inaccuracy and the long-standing definitional problems that have never been overcome, it is clear that the world is still in the midst of a sweeping and profound urban transformation that is literally changing the face of the planet."*


[^urban-def]: See for example @DepartmentforCommunitiesandLocalGovernment2006. This point is considered in [TODO: cross-ref].

```{r pop_data, fig.cap="Global population between 1950 and 2014, and forecasts up to 2050, according to \\citet{UN2014}\\footnote{The world's urban and total population forecasts can be found in Files 3 and 5, respectively}\\label{fig:pop_forecasts}", results="asis"}

## Raw data from https://esa.un.org/unpd/wup/CD-ROM/ (Files 3 and 5)
pop_data <- read.csv("01_introduction/UN-population-forecasts.csv")
pop_data <- melt(pop_data, id="Year", variable.name="Proportion")

pop_data$Year <- as.numeric(pop_data$Year)
pop_data$value <- pop_data$value / 1e6 ## Convert populations to billions
pop_data$Proportion <- factor(pop_data$Proportion, levels = rev(levels(pop_data$Proportion)))

pop_data_plot <- ggplot(pop_data, aes(Year, value)) +
  geom_line(aes(linetype=Proportion)) +
  ylab("Population [billions]") +
  theme_bw() +
  theme(axis.title.x = element_blank())
pop_data_plot
```

Urbanisation is the longterm result of humanity's search for the good life. People migrate from the countryside in search of employment, as well as better access to education, healthcare, and other services. Thus an urban population grows. As it does, firms benefit from economies of scale, and at the same time, cities start to become 'agglomeration economies' (in which different types of firms and industries benefit one another by way of supplying each other with goods and services, on account of their proximity to each other). With scale and agglomeration, cities produce specialised goods and services at low cost, which can be traded. Trade increases prosperity, and prosperity attracts migrants. Thus urbanisation becomes a virtuous cycle in which city size and economic growth feed one another (Figure \ref{fig:urbanCycle}) [@WorldBank2009].

\begin{figure}[h]
	\centering
	\input{01_introduction/urbanCycle}
	\caption{Virtuous circle of urban growth and economic prosperity.} \label{fig:urbanCycle}
\end{figure}

Urbanisation is a proven path to prosperity, so it is unsurprising that expert opinion reveals little appetite to oppose it. This is not to say they provide an immediate solution to all issues of poverty. Cities often have a widespread presence of slums, especially at the beginning of their development. The World Bank describes these as 'growing pains' [@WorldBank2009, p.69] and notes their historical precedence in cities like London, where slums are no longer to be found. The majority of growth takes place in the developing world. (Africa and Asia will together account for 86 per cent of the increase in urban dwellers by 2050 [@York2011]). It is encouraging that the urbanisation patterns seen historically in OEDC countries are being replicated in the developing world.
<!--
This growth is also Secondly, despite much excitement about megacities, urban growth is mainly in smaller cities (below 500,000 people) in the foreseeable future). Cities of more than 10 million residents will accommodate less than 10\% of the urban population [@Cohen2006]
-->

The number of cities, and the sizes of their populations are growing, driven by urbanisation's wealth-generating functions. While being home to over 50 per cent of the global population, cities generate around 80 per cent of global GDP [@WorldBank2016]. With urbanisation's benefits obvious, it is tempting to ignore the challenges it throws up. Some of these challenges -- in particular, those related to urban resource management -- which are introduced in the next two sections.

### Environmental challenges

There are many negative environmental consequences which can be traced to urban resource consumption. The environment is a huge field of study, and to summarise all negative environmental effects here would be impossible. However, just a few examples is enough to build make a compelling case that the levels of resource use by urban areas is environmentally problematic enough to warrant attention from researches, planners, and policy makers.

#### Sectoral strains

As a way in to give some sense of order to the environmental challenges, they will be listed in relation to the resource 'sector' from which they originate. This understanding of notion of 'sectors' will also become important an theme of the thesis[^sectors].

[^sectors]: The energy, waste, and water sectors have been chosen here since they will considered throughout the rest of this thesis, for reasons given in Section [TODO:]. Other sectors could have included transport, and materials (though these are arguably closely related to the energy and waste sectors).

<!--
Some environmental challenges may be manifested at easily identifiable locations (such as a city itself or its immediate surroundings), while other effects may be felt more generally. These include land-use impacts, resource scarcity, poor water quality, emissions, the heat-island effect and eutrophication, much of which can be related to resource overconsumption \citep{Cao2011}. Such negative effects can be seen in relation to three sectors.
-->

\begin{description}

\item [Energy] Cities are home to just over 50 per cent of the globabl population, yet they are responsible for around 60 per cent of primary energy consumption, and 71 per cent of energy-related greenhouse gas emissions \cite{IEA2008}. Climate change is a global problem (i.e. it does not necessarily affect cities in proportion to their emissions. Thus rightly managing the urban energy sector can rightly be considered a matter of global importance.
%There are multiple causes of this including the lifecycle heating and power demands, and urban sprawl leading to increases in transport use due to longer and more frequent journeys [@Grubler2009].

\item [Waste] The primary issue here is simply the volume of waste produced, which municipal governments need to deal with somehow. With their higher quality of life, urban dwellers consume more goods, and so generate more waste. Thus per capita waste generation rates are higher in cities than in rural areas. With the world both urbanising, and those urban dwellers continuing to improve their living standards, the World Bank estimates that total municipal solid waste could increase by 70 per cent between 2012 and 2025 \citep{Hoornweg2012}.

Other waste-related challenges include urban waste's contribution to climate change -- about 5 per cent of total greenhouse emissions come from waste, as the combined contributions of methane (from organic waste) and carbon dioxide (from incineration, which also leads to air pollution) \citep{Hoornweg2012}; flooding due to the blocking of drainage \citep{Lamond2012}; and public health concerns \citep{Kouame2014}, bad odours, and aesthetic impacts \citep{Shah2012}.

%These include groundwater contamination, public opposition to waste disposal near residential areas [@Li2006], as well as growing waste generation rates and decreasing disposal capacity [@Lu2009]. For example, [@Xydis2012] have described the waste management capacity limitation in Greece as having reached a 'critical point'. The decomposition of waste in landfill sites contributes to global warming through the production of methane (in addition to the emissions from the transport and management of waste [@ICE2011]). The health risks posed by open sites in developing countries are also of concern [@Durand2013].

\item [Water] While there is a problem of overconsumption in this sector, it is of a different nature than the equivalent energy- and waste-related challenges, as waster is a resource which does not have a problem of accumulation (like emissions or waste). Rather, water is cycled around the planet and the challenge is to get water in the right place at the right time, treated to an appropriate standard. As urbanisation increases affluence, and business and industrial activity, per capita water consumption rises \citep{AlFry2006}. Water stress (where the available water cannot meet all uses) is a commonplace problem -- about half of the world's cities with with populations greater than 100,000 cannot replenish their water resources at the rate they use them \citep{Richter2013}.

The cycling behaviour of water means the effects of water stress may not be felt where they are caused, since the management infrastructure (for treatment and transport) can be afforded by the affluent, but not by the poor (thus water stress is disproportionately suffered in the developing countries, which is also where most urbanisation is happening). Thus many large cities use networks extending up to 200 kilometers over their surrounding area to meet demand, but for many urbanising regions, it is difficult to invest quickly enough to build the infrastructure to meet demand \citep{EEA2016}.

\end{description}

With just a brief look at each of the above sectors, it is clear from the expert opinion that there are serious concerns to be addressed. The common cause of these problems is that of affluence. Urban areas have proved successful at providing a high quality of life, but this is inevitably in tension with the detrimental effects on quality of life which will emerge.

The backdrop against which many of the above issues are set is climate change, which is arguably the greatest environmental threat of all [@WEF2016]. With the energy and waste sectors large contributers of greenhouse gas emissions, coupled with the fact that climate change limits the availability of water in many places [@AlFry2006], climate change links the challenges of the energy, water, and waste sectors.

#### Intersectoral interactions

Not only do the energy, water and waste sectors all interact with climate change, they also interact with each other, such that issues of once sector feed the issues of another, thus exacerbating the environmental challenges. These intersectoral interactions are introduced briefly, since this will become an important theme in this thesis, and are summarised in Figure [TODO: cross-ref] 

\begin{description}

\item [Energy and water] As demand for water increases, so does the need for the energy required for its treatment and distribution. It is estimated that approximately 35 per cent of the energy used by municipalities in the United States is for water supply and wastewater treatment \citep{Elliott2005}. 

Climate change serves only to this energy-water nexus challenge more difficult, since as it limits the availability of water in many places, thus increasing energy requirements (e.g. to provide water via desalination and other energy-intensive treatment methods). Of course, the increased energy use further exacerbates climate change. Thus the energy-water intersectoral interactions become a self-reinforcing challenge. \citep{Webber2011}

In the other direction, energy use increases water consumption through the need for cooling water in power plants \citep{McMahon2011}. 

\item [Energy and waste] The waste sector has a large energy consumption corresponding to transportation needs (mainly vehicle) and management processes (such as sorting, recycling, and crushing). \citep{ICE2011, Hoornweg2012} 

In the other direction, increased energy consumption increases waste output, for example, the amount of fly ash produced by coal plants in the United States is over half of the quantity of municipal solid waste\footnote{Annual MSW generation as about 254 millions tons \citep{EPA2014}, and annual fly ash generations is about 130 million tons \citep{WorseThanNuclear2012}}. Radioactive waste from nuclear energy generation is another significant challenge \citep{NEI2016}.

\item [Water and waste] Increases in water consumption leads to increased wastewater generation which needs to be managed through either treatment or disposal \citep{McMahon2011}.

In the other direction, wastewater is often discharged into freshwater systems. While nature can purify and recycle contaminated water to some extent, this is happening at a rate that nature cannot keep up with (and so expensive treatments are required) \citep{AlFry2006}. Solid waste is also problematic in this regard because leachate contaminates groundwater and surface water sources \citep{Hoornweg2012}.

\end{description}

<!--
https://www.theguardian.com/global-development/2012/jun/06/world-urban-waste-mountain
- "The report notes that municipal solid waste management is the most important service a city provides. In poorer countries, rubbish collection and processing is often the largest single budget item for cities, and one of the largest employers."
- "A city that cannot effectively manage its waste is rarely able to manage more complex services such as health, education, or transportation, according to the report, and improving waste management is one of the most effective ways of strengthening overall municipal management."
-->

<!--
[TODO: MOVE]
Another conclusion that is inevitable from this survey of challanges is that resource efficiency alone isn't enough. The world is already struggling to sustain the global population (for example, at current trends, the entire global biomass would be required to meet energy needs by 2050). Thus, a 'per capita' decrease in consumption in a growing population will not be sufficient to meet the challenge. Moreover, more efficient use of resources may cause people to seek a higher standard of living -- the so called 'take-back' effect. There must be dematerialisation in absolute terms [@Winiwarter2011].
-->

### Economic stability

The way that cities use resources does not just present environmental challenges, but economic challenges too (perhaps less familiar). Briefly, it is possible to show that a city's population can grow at such a rate that is impossible to sustain with the available resources [@Bettencourt2007a].

This can be shown from equating a couple of equations. First, a now well-established idea is that at time $t$, quantities in cities $Y(t)$ (such as energy consumption or wealth), scale with population, $N(t)$, according to the following power law:
<!--
As entities that both consume and produce resources, they are analogous to living organisms and ecosystems, whose properties (such as body size or growth rate) scale in a 'self-similar' manner with mass (obeying Zipf's power law). In the case of a city, indicators, $Y$ (such as GDP or energy consumption) scale with population, $N$ as:
-->

\begin{align} \label{eq:urban_scale}
	Y(t)=Y_0N(t)^{\beta}
\end{align}

where $Y_0$ is a normalisation constant and $\beta$ is a constant which corresponds to the indicator under consideration. This result is interesting in and of itself, as it reveals that cities of different sizes essentially scaled versions of one another, and thus different cities do in fact operate in very similar ways [^animal-analogy], such that it is possible to predict many different quantities from just knowing the population.

Even more interesting are the values which $\beta$ takes for different types of $Y$ quantities. It turns out that for material quantities (such as the number of petrol stations, or the length of electrical cable), $\beta<1$; and for social quantities (such as GDP, or the number of serious crimes), $\beta>1$. In the former case, scaling is sublinear, and quantities exhibit economies of scale, such that bigger cities can serve their cities more efficiently (e.g. the number of gas stations required does not need to increase as quickly as the population). In the latter case, scaling is superlinear, such that bigger cities benefit from agglomeration effects of knowledge spillover and other social dynamics (e.g. wealth increases more quickly than that population)[^animal-analogy-departure]. This tension will become important later.

[^animal-analogy]: In this way, cities are analogous to animals, whose physiological characteristics (e.g. metabolic rate) scale with its mass.

[^animal-analogy-departure]: It is here that city behaviour departs from that of animals. In their case, all scaling is sublinear, i.e. biological systems develop to optimise their space, and thus behave according to economies of scale. However, there is nothing in animals that corresponds to the social characteristics of a city.

Second, population growth can be expressed as a function of the resources required to sustain it. If each individual requires some quantity, $R$, to maintain them per unit time, and $E$ (extra) to accommodate a new member, then that total total resource requirement, $Y$ is given by

\begin{align} \label{eq:urban_growth}
	Y(t)=RN(t)+E\frac{dN(t)}{dt}
\end{align}

Equating Equations \eqref{eq:urban_scale} and \eqref{eq:urban_growth}, and solving for $N(t)$ gives
<!--
\begin{align*}
	\frac{dN(t)}{dt}=\left(\frac{Y_0}{E}\right)N(t)^{\beta}-\left(\frac{R}{E}\right)N(t),
\end{align*}
-->

\begin{align*}
	N(t)=\left\{\frac{Y_0}{R}+\left(N^{1-\beta}(0)-\frac{Y_0}{R}\right)\exp\left[-\frac{R}{E}(1-\beta)t\right]\right\}^{\frac{1}{1-\beta}}
\end{align*}

<!--
Thus material economies of scale is in tension with the super-linear scaling of innovation characteristics. 
-->

Of pertinence here is how $N(t)$ grows in cities where agglomeration effects drive the population growth, $\beta>1$. The population will grow at above-exponential rates, theoretically reaching infinity after critical time, $t=t_c$\footnote{This crictical time, $t_c$ is given by \begin{align*} $t_c&=-\frac{E}{(\beta-1)R}\ln\left[1-\frac{R}{Y_0}N^{1-\beta}(0)\right] \\ &\approx \left[\frac{E}{(\beta-1)R}\right]\frac{1}{N^{\beta-1}(0)} \end{align*}}. In other words, in these cities, the social dynamics (e.g. the number of patents registered) move faster, but this same dynamic leads to unbounded growth (http://www.nature.com/nature/journal/v467/n7318/pdf/467912a.pdf). 

The problem of economic stability comes because in reality, resources $R$ and $E$ are limited, this unbounded growth cannot be sustained. At this point, $N$ stops growing and starts to collapse. To avoid this, requires innovation, to sustain agglomeration-driven growth. On @Bettencourt2007's model, this corresponds to resetting the values of Equation \eqref{eq:urban_growth}'s parameters, before $t_c$ is reached.

However, $t_c$ for each new cycle of innovation becomes shorter and shorter requiring innovation to take place at increasingly faster paces (see Figure~\ref{fig:bettencourt_innovation}). This need for innovation to happen after ever-decreasing time intervals is consistent with historical movements between the Stone, Bronze, and Iron Ages, and the Computer, and Information and Digital Ages [TODO: cite http://www.nature.com/nature/journal/v467/n7318/pdf/467912a.pdf].

<!--
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.75\textwidth]{01_introduction/bettencourt_growthPlots.pdf} 
	\caption{Urban growth regimes taken from \citet{Bettencourt2007}. (a) Shows sublinear growth. (b) Shows exponential growth (linear). (c) Shows superlinear scaling before resources become limited. (d) Shows the economic collapse in the superlinear regime when resources become limited.} \label{fig:bettencourt_growth}
\end{figure}

These regimes are plotted in Figure~\ref{fig:bettencourt_growth}.
-->

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.75\textwidth]{01_introduction/bettencourt_innovation.pdf} 
	\caption{Representation of the innovation cycles necessary to postpone economic collapse taken from \citet{Bettencourt2007}. Note how the critical time $t_c$ reduces for each cycle.} \label{fig:bettencourt_innovation}
\end{figure}

### Summary

Thus far, we've seen that urbanisation is helping more and more people live better quality of life, but the resource use which sustains these lifestyles is a cause of serious environmental and economic problems. Taking these issues together, the challenge can be considered one of *sustainable development*, first defined by the Brutland Commission as

> "...meeting the needs of the present without the ability of future generations to meet their own needs." [@Brutland1987, TODO: page number]

From the previous two sections, it is clear that 

- economic felt within a city
- environmental strain effects distributed more widely (less possible to calculate) -- i.e. planetary boundary issues

The sustainability issue is one for both cities themselves to consider, and for the world to consider as a whole, and this is manifested in various policy proposals. Using political instruments to address the global climate issues is familiar territory [TODO: complete]. Over recent decades, a united will to stop damage to the ozone layer meant [TODO: N] countries banned CFCs with the result that [TODO: continue]. In more recent decades, climate change has been a constant target of policy, with nations coming together at Kyoto and Paris [TODO: provide further details] to set emissions targets [TODO: etc.].

Less familiar are the policies concerning the other environmental issues outlined above, probably in part due to their specificity. Thus, as with the issues themselves, the policies proposed to address them are too numerous to list here. Around the world policies will vary according to need and political will; there is not the sense of global cause there is with the ozone layer or climate change. Nevertheless, policies do exist. This section gathers together just a few key voices from research and other areas, along with some specific policy examples to demonstrate that there is, at least to some extent, both a political will and a plan, to address some of these issues.

- Waste policy (World Bank): http://documents.worldbank.org/curated/en/302341468126264791/pdf/68135-REVISED-What-a-Waste-2012-Final-updated.pdf
- EU Policy Making (European Environment Agency): file:///home/tr608/Downloads/Urban%20sustainability%20issues%20-%20Enabling%20resource-efficient%20cities.pdf
- German example http://www2.giz.de/wbf/4tDx9kw63gma/FactSheet_09_Nexus_engl_21052013.pdf


Given the environmental challenges, it is no surprise that there is an increasing emphasis on developing policies and decisions to tackle the issues presented by urbanisation. For example, @Agudelo-Vera2011] argues that the objective of urban planning needs to move on from simply solving the problems of accommodating larger populations and meeting transport needs, to include sustainable development. The authors contend that resource management is an essential part of sustainable development, and that resource management should move on from its historic aims of simply meeting demand. Thus, the discipline of urban planning should aim to end the pattern of over consumption and excessive waste production which the global ecosystem cannot carry. This approach is manifested in policies such as the European Union target to provide 20\% of final energy from renewable sources by 2020.

To bring together the interrelated resource management challenges and policies, the city of London is considered, because it is a good example of a city experiencing resource management problems and thus adopting policies and actions to meet the challenges.

1. In the *energy* sector, London has a comprehensive range of policies concerning hybrid-fuelled transport, the retrofitting of housing and the decentralisation of energy supply in order to reduce energy-related emissions [@Strategy2011]. 
2. In the *water* sector, London is under large stress, suffering from an old network which leaks about 25% of the water which enters it. When this is coupled with increasing demand (as population grows, with a tendency to live in houses of fewer people) and the reduced availability of water due to climate change, then it is obvious that current consumption (already above the national average) is unsustainable. Thus policies are being implemented to capture rainwater and wastewater as a resource; change consumer behaviour and payment patterns; and make buildings more efficient [@Nickson2011]. 
3. In the municipal *waste* sector, 49% currently ends up in landfill, some of this outside London. Policies are being implemented to build new waste management infrastructure, increase recycling and use waste incineration as a method of energy production, for example with the SELCHP\footnote{South East London Combined Heat and Power.} incinerator [@Zabala2011].

<!--
\citep{Chen2006} Link to resource use, environment, policy and technology options.
\citep{Batt2010} to summarise.
-->

In summary, this thesis is in agreement with @Newman1999 that the environmental strain means that cities must take responsibility to manage their resources better.

Regarding the economic stability of cities, the policy environment is even less established. Perhaps this is because the main need is for that of 'innovation', which is a somewhat organic phenomenon -- it is not possible to set targets as for emissions or landfill diversion rates. Nevertheless, in our times, there is much study of and thinking about cities; how they behave, and how they should develop. Thus Bettencourt's work leads him to use a *Nature* comment piece to call for urban development policies which consider the growth dynamics he has highlighted, commenting that "the difference between 'policy as usual' and policy led by a new quantitative understanding of cities may well be the choice between creating a "planet of slums" or finally achieving a sustainable, creative, prosperous, urbanized world expressing the best of the human spirit." [TODO: http://www.nature.com/nature/journal/v467/n7318/pdf/467912a.pdf, page 913].

As an alternative to 'innovation', efficient resource management should be adopted in order to effectively increase the value of $R$ and hence buys innovation time to push the stagnation point into the future.

In addition to the policy proposals above, there is another concept -- key to this thesis -- that could be adopted to help meet these challenges. That is the properties inherent to a city which means they can manage resources efficiently.

## Urbanisation: the opportunity
\label{sec:urbanOpps}

The headline statistics show that compared to rural areas, that on a per capita basis, cities use more energy and water, and produce more waste. However, once quality of life is factored in, it is actually the case that, on the whole, cities use less resources per capita to provide a given quality of life. This is, in essence, a key principle of urban areas -- they can supply goods and services to their point of need more efficiently than smaller settlement types.

This efficiency must naturally follow from the scaling laws in @Bettencourt2010. If employment, GDP, and other quality-of-life indicators increase at a rate quicker than population growth[^qual-of-life], yet household consumption of energy and water increases at slower rate (see $\beta$ values in Table [TODO: cite table], then it necessarily follows that urbanisation is an efficiency-finding process. Even if one were to take overall electricity consumption (i.e. not just that of households) which grows at a quicker rate than the population, it is still a slower growth rate than GDP and employment.

There maybe detailed mathematical demonstration of this, in http://science.sciencemag.org/content/sci/340/6139/1438.full.pdf.

- Incremental network growth to serve population
- See Bettencourt's model of unitis of infrastructure at successive heirarchical levels ($h$) to show that land are per person, and shortes network deistance, decrease with N.
- Cities can be suboptimal for:
  	- failing to reach their social potential
  	- reaching social potential in a way that transport costs become too high
- acknowledges unequal distribution within a city
- infrastructure volume scales faster than land area (though both sublinear), e.g. underground
- Energy loss in transport processes follows socioeconomic scaling; unlike other complex systems. Implications for sustainable development
- Cities appear like networks found in physiology or natural geography, but they are more like stars -- "open-ended social reactors" (p.1441).

[^qual-of-life]: Note however, that negative indicators also follow greater-than-population growth rates, such as new AIDS cases ($\beta=1.23$) and serious crimes ($\beta=1.16$).

[TODO: tabulate a subset of beta values: GDP, employment, household electrical and water consumption, total electrical consumption. This table may be better off in the Economic stability section.]

- Note, there is variation among cities http://www.pnas.org/content/112/19/5985.full.pdf

### More efficient still?

This efficiency inherent to urban areas can largely be explained by one word -- *proximity*. The city faciliates the mixing of people the socioeconomic gains from this surpass the resource demands. However, there is another sort of proximity which applies to infrastructure, rather than people.

In cities, resource consumers (and producers) are close to the infrastructure that manages those resources. Here are a few ways this manifests in resource-efficiency. With resource-management infrastructure co-located in cities, and close to multiple types of users, there are opportunities to make further resource savings, using strategies increasingly adopted in cities.

\begin{description}

%\item [Waste collectors] In poorer areas, people see the opportunity in poorer areas (http://practicalaction.org/urban-waste-management-3)
%- Integrated Sustainable Waste Management (http://documents.worldbank.org/curated/en/302341468126264791/pdf/68135-REVISED-What-a-Waste-2012-Final-updated.pdf)
%Proximity of the population to waste

\item [Decentralisation] is a term from the field of energy management, to refer to energy generation that takes place closer to the point of use than traditional centralised generation. In electricity networks, this can reduce distribution power losses (and hence emissions) \citep{Fleten2007}. Of even greater interest, is that decentralisation enables a diverse mix of energy-generation technologies to use play their in the energy distribution, including combined heat and power (CHP), photovoltaic systems (PV), small wind turbines, biogas digesters and other systems \citep{Ren2010}. This permits a variety of local resources to be used as energy sources (including renewables such as wind, solar and biomass, as well as traditional fuels such as coal and gas).

Modelling work \citep{Webber2011} and empirical data [TODO: cite example] suggest that there are environmental and financial benefits to integrating traditional centralised energy management with  technologies distributed across multiple locations, fed by local resources.

Decentralisation can also be applied to water systems [TODO: cite http://www.waterworld.com/articles/wwi/print/volume-18/issue-3/features/decentralised-infrastructure-saves-system-costs-produces-biogas-energy.html], by using infrastructure to harvest rainwater through roof collectors and permable pavements, and reclaim or reuse waste water at locations distrubuted around a city, even down at the household scale. Resolving the collection and management (and even treatment) of water to local levels reduces freshwater withdrawals and energy requirements (for treatment and transport) \citet{Daigger2009}.

\item [Cascading] is another principle to improve energy system efficiency. The classic example is CHP in which the waste heat from electricity generation is used in heating systems \citep{Grubler2009}. In energy markets, energy cascading will manifest as companies provide 'integrated energy services' (selling both heat and electricity to the consumer) \citep{Sugihara2004}.

Cascading can be adopted in urban water management\footnote{The term 'cascading' is not used in the literature, but the prinicple has similarities with energy cascading.}, when wastewater from one purpose is used in for another purpose which requires less stringent quality standards. A household example could be the the separation of potable and nonpotable water streams such that grey water from the washing could be used in toilet flushing or watering the plants [TODO: cite http://www.thegreenage.co.uk/tech/greywater-recycling/].

[TODO: Move to intersectoral synergies]

Biomass is another resource that can similarly be 'cascaded' in ''energetic recycling" using residues and wastes in energy provision, rather than appropriating more natural resources for the same purpose \citep{Haberl2001a}.

\end{description}

Both decentralisation and cascading are are possible because of infrastructure proximity. The examples listed above show how cities can get the most of a particular infrastructure system. However, is is possible to find even more efficiency from urban resource management systems?

<!--
\citep{Cai2011} Systematic and intensive municiple infrastructure e.g. integrated water systems
\citep{Clift2000} Systems approach

% Energy
%% Point out that there are inherent advantages to DER: price, reduces emissions, developing countires, fuel prices, low carbon construction
-->

#### Intersectoral synergies

A specific type of synergy that can stem from proximity is that of the different sectors' infrastructure to each other, such that the waste and by-products of one sector can become the inputs to another, i.e. cascading between the decentralised resource management processes of different sectors. This is what \citet{Leduc2010} refers to as the 'urban harvest' approach, which takes advantage of connectivity and proximity in cities to maximise the reuse of water, waste and other materials within an urban system, and the cascading of energy through multiple uses in order to 'close' cycles. Having considered the shape of the energy and water sectors in isolation, the possible cross-sectoral interactions and intersections the urban environment gives rise to are now considered.

\begin{description}

\item [Energy and water] Opportunities to find synergies between these two sectors include hydroelectric power which converts water's hydrostatic and dynamic energy into electrical energy; water source heatpumps, which heat energy stored in water for some useful purpose; and anearobic digestion of wastewater which converts organic content of wastewater to biogas, which can be used as a fuel.

An indirect saving can be made via wastewater reclamation, which has the side effect of reducing overall water system energy requirements (by reducing the need to treat and distribute freshwater from scratch) \citet{Lundin2000}.

\item [Energy and waste] Synergies between these two sectors are increasingly advocated. One example is waste incineration; not only does this save on landfill space, the heat produced can be usefully diverted in either heating or power generation.

Anaerobic digestion can also be applied here, for waste with organic content (e.g. foodwaste) to provide energy \citet{Iacovidou2012}. 

Indirectly, any other ways of reusing or recycling waste will reduce energy required to manufacture new goods \citep{Geng2010}. 

\item [Water and waste]

[TODO: examples]

\end{description}

<!--
%% INCLUDE FIG 5 and 6. 
%%challenge of bioenergy increase in water use. (McMahon).
%%\citep{Lim2010} Heirarchical water allocation, integration of infrastructure, reducing electricity consumption from pumping etc.
%% footnote on how wastewater is classified is required.
%%\citep{Makropulos2008} Tool for IWM based on an holistic view. Tradeoffs
%%\citep{Nickson2011} Rainwater, waste-water (for energy) and reclaimed water (reduces pumping).
-->

<!--
%An underlying theoretical framework is required which can help us measure the scale of the challenges presented; the opportunities on offer; and the success of the models. The 'urban metabolism' concept will serve this purpose.
%
%Many studies suggest integrating infrastructures within their own field. This project is just an extension of that concept.
%
%The definition of integration (literature may be different from this study).
%The term `integration' can have multiple meanings
%Furthermore, there is little research in this field (despite many calls for it).  
-->

#### Summary

[TODO: diagram of links (note: only initial, but a systematic version will appear]

In summary, the co-location of water and energy infrastructure means synergies can be exploited in an 'integrated water system' such that both net water consumption and energy performance improves \citep{Makropoulos2008}.

Having seen how a more efficient use of resources is the route to meeting the problems caused by urbanisation, this section explores how urban areas are uniquely home to great opportunities to achieve this aim, chiefly due to the density of urban areas, and the resulting co-location of infrastructures. Thus, \emph{``urban planning in the 21\textsuperscript{st} Century should evolve towards an ecologically-oriented macro-architecture fully integrating the design and location of energy- and material-efficient buildings and urban infrastructure with overall spatial planning further to minimize material throughput.''} \citep{Rees1999}. 

The dense nature of cities is such that synergies between urban infrastructures can be identified and exploited. Thus a node in a resource management network can have multiple functionalities and the waste products of one process can become the inputs to another. Such 'circularisation' of resources is increasingly playing a part in sustainable urban design \citep{Meijer2011}. The rest of this section goes on to outline some of the specific resource interactions which exist in urban environments, considering the energy and water sectors in turn. From this starting point, the resource integration opportunities possible at the 'energy-water nexus' and the 'energy-waste nexus' can be outlined.

\item [`Industrial symbiosis'] is advocated whereby firms use each others' waste outputs as raw material inputs in other processes (e.g. production and manufacturing as well as WtE applications), in line with the 'proximity principle' observed in Japan, whereby waste is managed close to source. %For example, the OWARE model of \citep{Eriksson2002} takes a systems perspective on all the possible sources and uses of waste in a system (such as a city) by combining various sub-models which consider the life-cycle costs associated with incinceration, landfill, recyling , anaerobic digestion and so on.

Find its route to policy via urban metabolism [TODO: cite http://ac.els-cdn.com/S0921344912000250/1-s2.0-S0921344912000250-main.pdf?_tid=e4246f32-6bb1-11e6-8231-00000aab0f02&acdnat=1472232426_433302133d00ea76ce4dfa27702e30a7, page 4].

[TODO: summary text]

## Systems optimisation

Listing the intersectoral synergies is all very well. However, if planners and policy makers are to make the best use of these interactions [TODO:]. It is easy to see that it is not necessarily straightforward. Cities will vary in the precise demands they need (quantities and locations), as well as the infrastructure they have to meet those needs, and the ability they have to make decisions about that infrastructure.

But in practice, there needs to be a comprehensive understanding and application of them. One 

Returning to consider two themes introduced earlier

- the mechanism behind the success of the city (proximity/co-location)
- the concept of a 'system'

The possibility is there. The question is, how can it be implemented (at least the engineering aspects)? For the answer, we look to systems thinking.

In fact, one does not need to think about energy, water and waste sectors together to come across systems thinking.

The more we consider the intersectoral synergies, the more we will appreciate that resource management can be thought of as a 'system'. Multiple components (which include resources and the infrastructure which manage them) are related in some way.

Examples of individual sectors

A simple model to optimise the flows in an energy-waste network is proposed by \citet{Kharrazi2012} as an example of an integrated resource network. Nodes represent the producers and consumers of waste and electricity, and links between the nodes are the flows of those resources. A real world application of WtE networks is seen in Greece bringing two-fold benefits: mitigating against landfill shortages and producing renewable energy \citep{Xydis2012}.

Example of 'industrial symbiosis'

Brief summary of the mathematics/optimisation

Comment that the literature review will show the need for an integrated model.

The policy environment suggest there is at least to some extent a will and a plan for cities to manage their resources better. However, one part of the plan for which there is certainly a will, but perhaps less of a plan, is the idea that a city can be understood as a system, and interventions can be made which improve the overall performance of the system.

## Urban metabolism -- the theoretical framework

Examples of the explanatory power of UM [TODO:], and indeed, most of what has been discussed thus far can be explained with reference to UM.

- In general, urban areas both experience an internal environmental strain and exert an environmental strain externally. A society starts to overconsume resources when the area which serves it (the hinterland) is unable to provide enough food, water and other materials to meet the needs of its inhabitants. An increasing population requires a larger quantity of resources to meet its needs. To extract and manage these resources itself requires a growing population inducing a self-feeding cycle of population  growth and resource management. At some point, a 'labour limit' is reached where the land can no longer sustain the workforce. Such resource scarcity can be true of both agricultural based societies as well as industrial based societies [@Haberl2001a, Haberl2001b, Gr2003]. In seeking to overcome these internal pressures on resources, cities look beyond the immediate hinterlands, making use of the transport system to become dependent on resource imports from global markets. Thus the environmental impact that cities are responsible for spreads such that it becomes an external, global problem [@Agudelo-Vera2011]. In summary, the internal environmental strain due to resource over-consumption leads to external impacts on a global scale.
- Urbanisation is leading to 'water stress' as populations grow and there is a demand for a higher quality of life. One example of how this can happen in urban areas is given by \citet{Kennedy2008} who considers the relationship between urbanisation and groundwater. At the early stages of urban development, shallow wells are sufficient to supply a population, and wastewater can discharge to a watercourse. As the population grows and water extraction rates rise, the water table falls requiring the digging of deeper wells. Shallow groundwater can also become contaminated by wastewater discharge. Thus water must be sourced (at significant expense) from outside the city, changing the city's groundwater characteristics such that the water table can rise above its original level and even cause flooding. In summary then, water security issues in urban areas are causing financial, health and environmental concerns.
%Associated with the increasing demand is the need to treat the water and prevent contamination. This demand for water in high quantities comes with a financial burden [@Daigger2009]. 

The environmental and economic issues above can be brought together under the notion of 'sustainability'. [TODO: comment on the triple bottom line] 

Closely related to the field of urban sustainability is the concept of an urban area's 'metabolism'.

Perhaps one unexplored area of urban metabolism is the idea of systems optimisation [TODO: complete] 

Thus the challenge tackled in this thesis can be framed in terms of urban metabolism (UM)


- thus, the notion of 'overconsumption' can be considered as when cities rely on their hinterlands and global markets (e.g. Haberl)?

The field under which urban resource consumption has been studied is urban metabolism. [TODO: brief definition]. By analogy with the body and eco-systems. Thus the challenges tackled in this thesis can be framed in these terms. How can a city improve its metabolism? The lit-review will introduce UM in some detail, but for now: [TODO: necessary developments to the field] 

Studying urban resource management is potentially a fragmented collection of study areas (where different resources are siloed into separate fields, where consumption is seen in isolation to urban form and functions, where [TODO: etc.]). So it is helpful that a field has grown up which to some extent merges these considerations. Thus UM studies can answer questions like:

- How have patterns of resource use changed for city A?
- How does the function of the city affect its metabolism?
- How does the maturity of the city affect its metabolism?
- What problems are caused by this city's particular resource use patterns?
- To what extent does a city have linear or circular resource management patterns?

But 

n conclusion, there are two major challenges that urban planning must face. Firstly that of using resources more efficiently in order to minimise environmental strain experienced in and caused by cities. Secondly, in economies where innovation is the primary driver of growth, the economic collapse that can occur as cities outgrow their resource feed needs to be tackled by more efficient resource management in order to buy innovation time. In both cases, the root cause of the problems is an overconsumption of resources, and thus the solution is found in consuming fewer resources overall.

It has been argued that cities are overconsuming resources, yet that they are also home to the opportunities for more efficient resource management. An underlying theoretical concept is required which can quantify the efficiency of urban resource use. To this end, the 'urban metabolism' (UM) concept is introduced. Thus, the discussion surrounding the linearisation and circularisation of resources is actually a comment on the nature of the metabolism of an environment. UM is useful in three ways:

1. UM studies provide data which \emph{motivates} the development of resource integration models. 
2. It quantitatively demonstrates the \emph{means} by which urban resource management can be improved: namely, a move away from resource 'linearisation' towards 'circularisation'. 
3. It provides a way of \emph{measuring} the success of the models. 

In this section, UM will be defined to show how it indicates and quantifies the \emph{means} of resource integration. Then UM studies will be cited demonstrating such means by which cities can meet challenges and opportunities.

Conducting a UM study can be used to understand the processes taking place within it. Such studies imagine the city as an organism (or an ecosystem) and aim to quantify the fluxes of water, energy, waste and other materials into and out of urban populations in order to find the \emph{``sum total of the technical and socioeconomic processes that occur in cities, resulting in growth, production of energy and elimination of waste''} \citep[p. 44]{Kennedy2008}. It is a broad discipline, and as such individual studies vary in their scope, about the inclusion of particular resources, and whether the study is related to socioeconomic or political concerns (as in \citet{Hobbes2007}). The key features of a UM study are now described.

1. The basic \emph{methdology} in a UM study is a 'material flow analysis' (MFA) where resource flows in and out of areas are measured to see where they accumulate \citep{Barles2009}. Inputs included 'domestic extraction' (DE) or `imports' whilst outpus are 'exports', `deliberate disposal' (DD) or 'wastes and emissions' (WE). 
2. The basic \emph{results} of an UM study are 'indicators' given by operations on these variables. These can be used to compare the metabolism of different locations (or the same location at different times) to assess the relative sustainability of an area\footnote{For example, $\mbox{`material intensity'}=\mbox{Direct material input}/\mbox{GDP}$. This value reduces as a society dematerialises.} \citep{Hobbes2005}. \citet{Haberl2001a, Haberl2001b} show that the indicator 'human appropriation of net primary production' (HANNP), 'domestic consumption' and 'total energy input' divides societies into three categories: hunter-gather, agricultural and industrial. Each of these exhibits qualitatively different metabolic profiles. Hunter-gather communities have no struggle meeting their needs from their environment, however agricultural communities will reach a 'labour' limit where the environment will not be able to meet the needs of those working on it (see Section~\ref{sec:urban_challenges}), because their energetic needs (including that embodied in food and materials) simply cannot be met by the 'biological productivity' of the area---thus even agricultural regimes can be unsustainable. 

Thus, UM is an accounting tool which offers a way to quantify the sustainability challenges presented by the overconsumption of resources. Hence it can be a useful consideration in sustainable urban design. \citet{Barles2010} makes the case for its importance in sustainable urban development issues, noting how research has moved on from viewing cities as .parasites', feeding off the environment at its expense. Instead, the UM approach has allowed the consideration of optimisation, making use of possible industrial symbioses, thus moving from linear to circular resource management in cities. Some specific results and features of UM studies which are pertinent to resource management efficiency are now discussed.

1. \emph{In general, there is an increasing per capita metabolism.} \citet{Kennedy2008} demonstrates the power of the UM concept in reviewing the studies of eight regions around the world since 1965. The authors conclude that in general, there is an increasing per capita metabolism for water, wastewater, energy and materials (therefore cities are assimilating mroe of these resources per person). The largest components in UM studies tend to be energy, water and waste, thus justifying the approach of this study in targetting these resources for resource integration opportunities. London is a good example of this result \citep{BFF2002}.
2. \emph{Improvements in metabolism are observed in some areas.} Despite the global trends, some places (such as Toronto) have seen per capita reductions in energy and water use. This has been achieved through increased efficiency. Similarly, multiple cities have increased recycling rates and hence reduced per capita waste consumption \citep{Kennedy2008}. 
3. \emph{High metabolism and economic indicators are linked}. A particular challenge noted in the literature is the coupling between direct material consumption (DMC), direct material inputs (DMI) and GDP, for example, in the case of Singapore \citep{Schulz2007a}. Thus \citet{Kowalski1997} argues the challenge that sustainable urban design faces is the need to de-link prosperity and quality of life from metabolism. This can be achieved as technologies increase in efficiency faster than the economic growth of a society (as well as bringing about changes in cultural expectations and behaviours). The way that such efficiency improvements can be achieved at the urban scale (as opposed to the individual technology scale) is through the circularisation of urban resource flows (rather than the current linear approach)---which ultimately reduces inputs, as argued by \citet{Newman1999} who advocates the 'extended metabolism model'. This approach wants to recognise that a city is more than a mechanism for the physical processing of resources, but also the centre of human opportunity. 

The results of these UM studies give the \emph{means} to achieving improved resource efficiency by identifying specific areas for circularisation and dematerialisation. For example, a multi-scale UM study of Paris and the surrounding region in \citet{Barles2009} shows the city's dependence on a large area for its waste management. Thus, actions can be targeted in that sector to improve Paris's sustainability. Having an appreciation for the utility of the UM concept in assessing the resource efficiency of an area, it can be seen why \citet{Kennedy2011} advocates UM studies in design, using MFAs to produce \emph{``more ecologically sensitive designs''} of a city through infrastructure integration, closing loops to minimise overall inputs and outputs of resources. In conclusion, it can be said that the UM concept is key to improving resource management in order to bring sustainable design considerations to urban planning \citep{Agudelo-Vera2011}.

<!--
%% definitions
%% challenges
% \citep{Chen2012} Measure of metabolism given by network environ theory. Indicators for linkage and synergism.
%% Opportunities
%% Policy and planning
% \citep{Hobbes2007} social context to relate UM to actors, decisions. Explanatory approach to MFA.
-->

## Aims and scope of the study
\label{sec:aims}

The subject of this thesis is an interesting topic of study. It requires both a broad understanding of the big picture of urban resource management in three areas (energy, water and waste), including knowledge of policy environments etc. coupled with an understanding of mathematical programming for optimsation models.

This Introduction has not attempted to be systematic in identifying all the challenges and opportunities in urban resource management. Nevertheless, there is enough here to suggest that there are environmental and economic issues to be tackled; and that there are enough intersectoral synergies, that they could become a viable source of urban resource efficiency.

Having now outlined the context and motivation for the thesis, a research question is posed. This gives the study a clear focus, and provides a benchmark for assessing the success of the final outcome of the study.

The following question will be answered:

> *By how much can the metabolism of an urban area be improved by creating and implementing models which optimise the integrated provision of energy, water and waste?*

This question comprises three aims:

1. Research the motivation, opportunity and methods for such models for the literature review. This will bring together up-to-date information on urban trends; a review of urban resource interactions and infrastrucures; and knowledge from existing methods used in optimising energy, water and waste management.
2. Develop models to calculate the optimal transfer of resources through a network of processes such that demand for resources of required quality is met. The models must capture the complexity of considering multiple resources and their qualities whilst remaining tractable. Three models will be developed during the term of study:
	i. A prototype model based on a small subsistence-based community. There will be no spatial or temporal disaggregation in this model, because its primary purpose is to develop a methodology which can handle multiple resource types, each of which are associated with one or more quality parameters.
	ii. A model to optimise resource management in an urban development (likely to be part of a city in China). This will introduce spatial disaggregation into the model.
	iii. A water-based model to optimise the management of the energy-water nexus. Unlike the previous two models (which are focused on design), this model will consider system operation, and will therefore include temporal disaggregation.
3. Assess how well the models improve urban metabolism through highlighting technological opportunities and pathways, and as a consequence, how well they minimise environmental impacts and ensure economic stability for urban areas.

There are four (potentially) novel aspects to this research: 
1. Whilst there are tools which consider the optimisation of two resources (for example, the minimisation of energy requirements for meeting water demands in a town), there are no such tools which simultaneously optimise energy, water and waste management. 
2. Secondly, a new method will be required which incorporates resource quality requirements in a manner that makes the model computationally efficient and tractable. 
3. Thirdly, the application of optimisation methods to the field of urban metabolism is new ground, with the literature dominated by accounting studies.
4. Fourthly, the tools developed will be new to an audience of urban planners, policy makers and utility service companies.

### Challenges dealt with here

- Define optimal metabolism, among different resource types
- Introduce optimisation modelling to UM
- Formulate a tractable problem which can represent an urban system
- Data needs
- Transparency: on the basis that modelling is about "insight" rather than numbers (Hamming).
- Transparency/Open source: to assist with both the introduction of UM optimisation, and the data needs

### Challenges not dealt with here

There are many fascinating aspects to cities, some of which touch on the themes already introduced. However, 

- institutional and creative forces

- socioeconomic/political/agent-based (transfer to the real-world)
- system operation

It must be remembered however, that a city isn't simply a resource processing machine, but it is the home to many people, and the place where they seek economic opportunity.  Therefore, there is a particular challenge to better manage resources without impacting upon the 'livability' of cities.

### Summary and contributions

In conclusion, there is strong case to pursue research into a model that optimises the provision of urban energy, water and waste resources. This is because urbanisation has socioeconomic advantages, such that it should not be resisted, despite the environmental and economic challenges brought by resource overconsumption. Rather, urban centres provide an opportunity for more efficient resource management by exploiting the proximity and co-location of infrastructure. The urban metabolism concept can be used to quantify the resource efficiency of a city, hence providing an underlying theoretical framework to the project.%he challenges of resource management. On top of this there is a research-gap in the field of integrated resource provision.

Table of contributions

<!--
\subsection{Report structure}
%Having presented the background to the study to provide motivation for the research question posed, the report continues by reviewing existing methods of optimising the provision of energy, water and waste in Section~\ref{sec:methods}. In Section~\ref{sec:models}, prototype models at their initial stage are presented. The report conludes in Section~\ref{sec:conc} with preliminary results, and an outline of the work to be undertaken in the rest of the project.
%
%In summary, it has been argued cities need to focus on resource integration in order to reduce damage to the environment and economic risk. Currently, there are no models which consider the optimal management of multiple urban resources.
-->

### Methods and thesis structure

Note the methodological challenge of the literature review(s), and how we meet that challenge through online search tools combined with bash and R programming.
	- to gather data
	- to locate our work within its related fields

By the end of the literature review, we will be suitably placed to precisely define each term in our title, but for now, it roughly approximates to "some computational means to exploit synergies to meet demand for goods and services in cities".

Diagram of thesis outline.
