# Technology database chapter

\epigraph{``It is a capital mistake to theorize before one has data.''}{Sherlock Holmes in \emph{A Study in Scarlett} (Arthur Conan Doyle)}

[TODO: complete]

[TODO: add note on which paper this comes from.]

```{r Setup, include=FALSE, results="hide", warning=FALSE}
opts_chunk$set(fig.path = "figures/", dev=c("cairo_pdf"), fig.pos='htbp', fig.width=5, fig.height=1.6, dpi=700, fig.show='hold', fig.lp="fig:", cache=FALSE, par=TRUE, echo=FALSE, results="hide", message=FALSE, warning=FALSE)

library(extrafont)
font_import(pattern="LinBiolinum")
```

## Introduction 

In 1965, the American engineer Abel Wolman introduced 'urban metabolism'. Writing in *Scientific American* he defined the metabolic requirements of a city as "all the materials and commodities needed to sustain the city's inhabitants at home, at work and at play" [@Wolman1965, p.156]. He recognised that resource consumption was necessary to sustain urban living, but was motivated to study urban metabolism because of his concerns that as cities grow, their rising material and energy inputs and outputs were causing problems with water supply, sewage disposal and air pollution.

[TODO: copy following paragraph to thesis introduction?]

Since Wolman, urbanisation has continued apace with more than 50% of the global population currently living in towns or cities, and predictions that this will exceed 70% by 2030 [@UnitedNations2014]. Metabolic problems have also grown in both magnitude and variety. Urban energy consumption is estimated to contribute up to 76% of global CO$_2$ emissions [@IPCC2013]; urban water consumption rates make water shortage a perennial reality for 150 million urban inhabitants [@McDonald2011]; and urban waste output levels are leading to water contamination, air pollution, greenhouse gas emissions and public health concerns [@Hoornweg2012].

As the environmental problems associated with urbanisation have increased, so urban metabolism has become an important tool for developing sustainable cities [@NextCity2014]. Specifically, urban metabolism studies could provide insight on how to maintain the socioeconomic benefits of urbanisation (highlighted by the @WorldBank2009 and others) while limiting environmental harm. Studying it enables planners and policy makers to take stock of an area's resource consumption and output, ascertaining its long-term trends and comparing its own metabolism with other areas. The outcomes of studies can then inform policy analysis and urban design [@Kennedy2011].

Analysis can focus on the different parts of an urban metabolic system, each of which work together in a hierarchical framework as shown in Figure \ref{fig:metab}. At the bottom lie end consumers, namely the homes, business, industry and transport that exert demands for final energy and water consumption, as well as generating wastes. At the top are the aggregate exchanges between a city and the surrounding environment: both the import of natural resources required to meet consumer demand, and the export of a city's wastes and effluents. In the middle is a network of processes that link the top and the bottom, converting natural resources to demands, as well as treating and managing locally generated wastes. Typically, urban metabolism studies seek to quantify the overall metabolic flows at the top, and to locate the origins of these flows at their point of demand at the bottom.

In this paper we will extend published arguments that advocate giving greater attention to the middle of the system, in order to show how the mix of processes used to meet demand at the bottom will affect overall metabolic inputs and outputs at the top. To support such approaches, we identify the resource conversion processes that can shape urban metabolism both now and in the future, and quantify the resource consumption and production of each individual process, assembling this information into a database. We then discuss possible applications of the database, before summarising our work and suggesting areas of further work.

\begin{figure}
	\centering
	\begin{spacing}{1.0}
	\sffamily
	\resizebox{\columnwidth}{!}{\input{04_database/metab-hierarchy.pdf_tex}}
	\end{spacing}
	\caption{Three components to an urban metabolism system with example resource flows adapted from the hypothetical city in \cite{Wolman1965}.}
	\label{fig:metab}
\end{figure}

## Background

Our background discussion has two elements: first, we justify our proposal for a database of processes by reviewing current approaches to urban metabolism to highlight their merits and shortcomings. Second, we introduce two research areas which will facilitate the assembly and use of the database.

### Approaches to urban metabolism

#### "Top-down"

According to @Chrysoulakis2014, most urban metabolism studies take a top-down approach. Their aim is to quantify the impacts of an entire urban area's consumption and waste generation [@Sahely2003]. Typically studies do this using data aggregated at the city scale (i.e. it cannot be traced back to particular locations or activities) [@Chrysoulakis2014]. Several methods are described in @Sahely2003. One example is Wolman's seminal study which derived a hypothetical city's resource flows from national rates of consumption (water, food and fuel) and generation (sewage, waste and air pollutants). A more sophisticated approach might adjust these values; for example, a city's food intake could be calculated by adjusting the national average food consumption according to a city's grocery bill expenditure (as suggested in Sahely et al.). Ideally though, data would be supplied directly from appropriate organisations: water consumption data from the public or private providers; electricity consumption data from market operators; and waste data from municipal authorities. Once collected, the data can be used to analyse trends and make inter-city comparisons, to reveal how variables (such as climate, urban form, economic activity etc.) correlate with overall metabolic flows, as exemplified by @Kennedy2015's study of megacities. Top-down studies can inform policy, as seen in @Liang2011 who use government statistics in a top-down metabolism study of Suzhou, China. Their findings lead to precise proposals to achieve decarbonization, in this case focussing CO$_2$ mitigation strategies on vehicle pollution (because trends indicate that the city is on course for low-carbon development in other sectors).

#### "Bottom-up"

An alternative approach is to build up a picture of a city's metabolism from the bottom up, by quantifying resource flows at small scales before aggregating them over the region of study [@Chrysoulakis2009]. For example, @Kellett2013 describe a method to quantify a neighbourhood's CO$_2$ emissions by modelling the CO$_2$ flows in the neighbourhood's buildings, transport, humans, and vegetation components. For instance, to calculate the total emissions attributable to buildings, first building types are determined from architectural features (found from fieldwork, databases, and LiDAR information), enabling researchers to know the number of each building type of building in the neighbourhood; next the emissions for each type of building (such as an apartment) are estimated using a model which takes into account a building's occupancy, age and HVAC equipment. Finally, emissions for each building type can be multiplied by the prevalence of each building type to calculate total emissions.

Another bottom-up approach comes out of the Vulcan Project which measures CO$_2$ emissions at fine spatial and temporal resolutions, using networks of ground and air remote sensing technologies [@Asefi-Najafabady2014]. @Gurney2012 present results for the city of Indianapolis to understand the origins of its emissions, enabling decision makers to implement mitigation strategies, which are both effective and low cost.

Bottom-up methods often depend on high-tech tools including: sensor technology to measure energy, water, and pollutant fluxes; GIS and remote sensing devices to collect information on land-use, thermal radiance and rainfall; and advanced numerical modelling to simulate resource fluxes [@Chrysoulakis2009]. An alternative low-tech approach is to survey residents of a city about their resource use habits; for instance, deriving information on urban water flows by asking householders about their typical dishwasher loads and shower durations [@Eberlein2014].

#### Merits, limitations, and a complementary approach

Top-down approaches are methodologically simple, but they do not nurture an appreciation of the ultimate causes of emissions or other environmental impacts [@Kellett2013], and are thus unable to predict how metabolic flows might change under different scenarios, hence the power of bottom-up methods which can be used to test scenarios and policies. Kellett et al. exemplify this in their neighbourhood emissions model by altering land-use, building efficiency, transport fuels, and consumer behaviour to observe how overall emissions change.

@Zhang2013 notes that both bottom-up and top-down approaches fail to consider the mechanisms within the urban system which determine its behaviour -- or as we have denoted it -- the *middle* of the system. The top-down approaches are well suited to identify the scale of metabolic problems, and the bottom-up approaches contribute understanding on the ultimate causes of problematic flows, but appreciation of the middle will add to the insights of bottom-up studies by revealing the role of a the internal mechanisms between the top and the bottom of an urban system. Thus Zhang argues research should prioritise quantitative analysis of an urban system's components, which models urban systems as a network of nodes and flows. This representation of a system provides a framework to understand how fluxes circulate between nodes, thus enabling researchers to test how management interventions and other scenarios will change the distribution of fluxes between nodes.

If these recommendations are to be realised, the urban metabolism field should seek a comprehensive understanding of what constitutes the middle to enable modelling to take place; i.e. identifying and describing the nodes and the links within metabolic networks. From an engineered-system perspective (as opposed to say an economic one), these nodes and links correspond to processes and resource flows, respectively (as per part (B) of Figure \ref{fig:metab}). Studying nodes and links could suggest how a city's mix of processes could meet resource demand in cheaper and greener ways, beyond demand-side management. As a case in point, consider how a city meets domestic heating demand: bottom-up studies could explore how measures such as thermal insulation and lower thermostat settings reduce fuel requirements at the top. But studying the middle would explore how investment in processes could reduce fuel usage, for example showing how how heat demand could be met by waste heat from power generation rather than gas boilers. We brand this type of analysis 'process-oriented'.

### Supporting fields of research

#### Existing databases and technology scanning

We now consider three precedents for our proposed database to locate our work within the existing landscape of tools. The first example are the 'technology roadmaps' published by the International Energy Agency [@IEA2015a]. These constitute comprehensive reports which outline important energy management technologies in view of climate change targets. For example, the technology roadmap for hydrogen and fuel cells outlines various hydrogen using systems (e.g. fuel cell vehicles) and hydrogen generating technologies (e.g. electrolysers), and includes narrative reporting which describes how the technologies work; tables of technical parameters such as process capacity, efficiency, capital cost, and lifetime expectancy; and the results of modelling which indicate the role hydrogen technologies could play in achieving targets [@IEA2015b]. 

Second, and also from the energy sector, is the Enipedia Wiki, which is a collaborative environment that collects information about energy technologies, infrastructure, resources and other related issues [@Davis2015]. For example one section of Enipedia assembles a database of energy storage technologies, which notes their maturity, possible applications, operating efficiency, and lifetime expectancy, alongside descriptions of how they work.

Our third example covers a broader range of technologies, namely the ecoinvent life cycle inventory database which specifies the material and energy inputs and outputs for processes within energy supply, transport, packaging, electronics and other areas. The database is usually used to assess the environmental impact of supply chains to produce a good or service [@Weidema2013].

These examples have many useful applications, but none of them are geared towards urban metabolism. For example, the IEA roadmaps and Enipedia are limited to energy management processes, and furthermore they do not record the non-energy flows associated with these processes, such as water and waste. ecoinvent is more comprehensive, in both the processes and flows recorded, but its focus is on product manufacturing supply chains, and thus it misses metabolic processes such as drinking water treatment and HVAC systems. Thus to facilitate process-oriented approaches to urban metabolism, a new database is needed.

To build the database, we will use principles from technology scanning, which can be defined as "a way of taking a creative look at the world of technological developments and the cultural, regulatory, and business environments in which they emerge" [@Edgetech]. In general, it is used to support corporate strategy, by enabling industry to keep up-to-date with the current technology landscape, or to scan the horizon to identify emerging technologies [@VanWyk1997] (though the specific aims and methods vary on a case-by-case basis).

As an example, in healthcare, researchers identified technologies which would improve clinical care in the most financially viable way through searching literature, databases and digital libraries, as well as consulting industrial experts [@AHRQ2015]. In another case study from the railroad industry, technology scanning was used to sift through many possible technologies to identify those which reduce costs or improve services,
in order to direct research and investment efforts [@Martland2002]. Martland et al.'s method involves first a "general search" to identify technologies, and second "technology mapping" which researches the performance and cost of each technology, and goes on to classify technologies according to maturity, functionality and other variables. Their scanning proved valuable as it uncovered technologies (such as nanotechnology) which had never before been identified in railroad technology strategies.

#### Open data

Open data is "data that can be freely used, re-used and redistributed by anyone -- subject only, at most, to the requirement to attribute and sharealike" [@OpenKnowledge2015], and is increasingly prevalent in academic research (Figure \ref{fig:open_data}). A high profile example is the Human Genome Project which placed all their human genomic sequence information in the public domain, so research would be of maximum societal benefit [@HUGO1996]. The term 'open data' first appeared in 1995, and was clarified in 2007 to incorporate three principles: "openness, participation and collaboration" [@Chignard2013]. Open data brings many benefits, including: opening up scientific enquiry, promoting new research, allowing verification of published results, and enabling research to go in directions not envisaged by the initial investigators [@Uhlir2007]. Added to this, making data open protects it from loss due to hardware malfunction, and also makes it a useful teaching tool [@Roche2014].

Considering open data within Industrial Ecology (IE), @Davis2010 argues that the community should improve its data management between sub-disciplines, so that research findings can more effectively contribute to the sum total of IE knowledge. @Zhu2014 offer an application of open data within IE which uses an open data set from the European Pollutant Release and Transfer Register (E-PRTR) [@EPRTR2009], to identify the type and location of pollutant releases from over 30,000 industrial facilities. They propose that this information be coupled with a dataset of industrial process inputs for European facilities, to matchmake facilities, so that the waste outputs of one facility meet the input requirements of another.

```{r open-data, fig.height=2, fig.cap="The growing research interest in open data since its first occurrence in 1995 until 2014. Data from the Scopus analysis tool, from searching titles, abstracts and keywords for '\\texttt{open data}' or '\\texttt{open-data}'. \\label{fig:open_data}"}

open_data <- read.csv("04_database/open-data-research.csv")

library(MASS)
rate <- fitdistr(open_data$Number, "exponential")

library(dplyr)
open_data <- filter(open_data, Year>=1995)

library(ggplot2)
plot_open_data <- ggplot(open_data, aes(Year, Number)) +
  theme_bw() +
  ylab('Number of publications') +
  theme(axis.title.x = element_blank()) +
  geom_point()
plot_open_data
```

## Assembling a database of resource conversion processes

There is a research need to study the middle of urban metabolic systems using process-oriented approaches. We will facilitate this by assembling an open database of as many processes as feasibly possible which could be defined as a node in the middle of an urban system. To do this we will adopt principles from technology scanning, however for our purposes we will rename this to 'process scanning' which is more appropriate for the larger-scale activities (which may involve more than one technology) that provide energy, water and waste management services in cities (such as a wastewater treatment plant). We adapt @Martland2002's two-stage methodology: in the general search stage, we systematically search academic literature to identify the processes; and in the process mapping stage, we collect data on each process and assemble the database.

### Systematic literature search (the general search)

Our systematic literature search uses online search engines, together with the Bash and R programming languages to identify relevant processes from the literature available online, following the procedure below:

1. *Assemble the titles of relevant journals.* We downloaded the titles of all the journals listed with the Thomson Reuters Master Journal List[^WebOfScienceURL] (at the time of download there were 17614 journals). To reduce this to relevant journals, we first carried out an automated search which removed journals that did not match any words from a specified list keywords[^GitHub-journals]. We subsequently reduced this to 379 journal titles with a manual search to remove the rest of the irrelevant content.
2. *Filter journals by impact.* This step reduces the literature to a manageable volume, which represents the highest quality research. First we removed any journals with impact factors less than 0.5. From the remaining journals, we retained the top 75% of the 'impact' by ranking the journals from highest to lowest according to their impact factor, and calculating a cumulative sum of their impact factors, and then removing all journals in the bottom 25%, leaving 168 journals.
3. *Find relevant journal articles.* Using the Web of Science search facility, we searched our list of journals for articles containing specified combinations of keywords[^GitHub-articles] which indicate they could include content related to urban resource management processes. This found 9929 paper titles which we reduced to 7681 titles through manual filtering.
4. *Article categorisation and further filtering by impact.* We then searched the article titles for keywords that enabled us to organise the articles into categories representing either a process or a resource. Each category was then divided into sub-categories. For example, the `effluent` category (a resource) contained sub-categories on `agriculture`, `fermentation`, `hydrogen` and `wastewater` amongst others. Each sub-category is therefore a list of journal articles whose titles contain both the word 'effluent' and the sub-category heading. This resulted in 65 main categories with a combined total of 955 sub-categories. Finally, we removed the lowest impact categories by calculating each category's mean article impact factor (totaling the impact factors of each article in a category before dividing by the number of articles in the category), then ranking the categories from highest to lowest mean impact, and removing categories in the bottom 33% of mean impact. This filtering left us with 4,747 unique article titles, thus making the process mapping stage manageable.

[^GitHub-journals]:\url{https://github.com/tomravalde/metabolism-database/tree/master/literature-search/keywords-journals.txt}
[^GitHub-articles]:\url{https://github.com/tomravalde/metabolism-database/tree/master/literature-search/keywords-articles.txt}

<!--
Terminal commands for getting the number of unique article title in the final step of the above process

for i in $(find . -name "*.csv" | grep -v processes); do echo $i; cat $i >> file.out; done
sort file.out > sorted.out
uniq sorted.out > uniqe.out
-->

[^WebOfScienceURL]: \url{http://science.thomsonreuters.com/cgi-bin/jrnlst/jlresults.cgi?PC=MASTER}

### Data collection and database assembly (process mapping)

Having categorised the processes and resources, we then manually searched each category for articles which refer to processes that manage energy, water or waste. For example, if an article referred to a fuel cell which treats landfill leachate, we would search for information to describe the leachate treatment process, using the original article, and if necessary other literature.

We used the YAML[^YAML] format to record the information for each process. YAML uses mappings to relate a value to a key. Mappings can be nested like sub-points and sub-sub points (and so on) in a bullet-point list by using a mapping of mappings, thus we can attribute multiple layers of information to any mapping. Moreover, YAML uses very lightweight syntax and is therefore human-readable when compared to other similarly structured data storage formats (such as JSON and XML). Furthermore, YAML data can easily be converted into other formats. The main keys making up a process's YAML file are as follows:

[^YAML]: 'YAML Ain't Markup Language': \url{yaml.org}

\begin{description}
\item[\texttt{process}] identifies the process name. For some processes, there exist a few variants, with different \texttt{capacity} values (see below); thus this key distinguishes these variants from one another by specifying both the process name (e.g. \texttt{anaerobic-digestion-wastewater}) and its capacity (e.g. \texttt{557} kg/s).
\item[\texttt{flow}] contains a list of mappings which specify the quantities of energy, water and waste consumed and produced by the process. Data on resource flow quantities come from process descriptions, inventories, and process mass and energy balances in the literature. Each flow contains another list of mappings comprising three items: the numerical quantity of the flow (which is negative for an input and positive for an output), the units (MW for energy quantities, and kg/s for everything else), and a reference to the information source.
\item[\texttt{trl}] defines the  `technology readiness level' (TRL) which is used to assess a technology's maturity, from the initial stages of scientific research to its implementation in the real world \citep{Mai2015}. TRLs were initially developed by NASA as a nine-point scale, but other organisations such as the U.S. Department of Defense use different scales \citep{Mankins2009}. We have defined our own simplified TRL definitions which define a process as `current' (9 on NASA's scale, for processes found commonly); `rare' (also 9 on NASA's scale, where processes exist in only a few places); `mid-term' (5--8 on NASA's scale, for processes which might have working pilots); or `long-term' (1--4 on NASA's scale, for those processes which only exist in the literature or the laboratory). For each process, we found this information by searching the literature and other online content. We used sub-mappings to record a justification of the TRL value attributed to a process, and a reference to the source of the justification.
\item[\texttt{main}] specifies which resource is considered the main input or output. For a power plant, this would be the electricity output; for a wastewater treatment plant, this would be the wastewater input. All the resource flow quantities are normalised with respect to this value, such that its magnitude equals 1.
\item[\texttt{capacity}] specifies the maximum rate at which the process can consume or produce the main resource. For example, for a 350 MW power plant, capacity would take a value of 350. Due to the normalisation of the flows with respect to the main resource, any process's resource flow multiplied by its capacity will provide the maximum rate of consumption or production of any resource for the process. This item contains two mappings specifying both a value and a reference.
\end{description}

The database is available as a GitHub repository at \url{https://github.com/tomravalde/metabolism-database}. An example database record is given in Figure \ref{fig:yaml}.

\begin{figure}
\lstinputlisting[language=yaml]{04_database/powerplant-paper-edit.yaml}
\caption{An example YAML record for a coal-fuelled powerplant.}
\label{fig:yaml}
\end{figure}

## Database overview and possible applications

```{r database-analysis}

# Import and shape data so we can obtain summary statistics

processes <- read.csv("04_database/processes-analysis.csv")

## Specifiy whether flows are inputs or outputs
processes$sign <- "Input"
processes[processes$value>0, ]$sign <- "Output"

## Add 'mainOutput' column

### Build mainOutput lookup table
lookup <- filter(processes, value==1)
lookup <- dplyr::select(lookup, c(process, resource))

### Use keyFlows.lookup to add 'mainOutput' column to proc.keyFlows
processes <- inner_join(processes, lookup, by="process")
colnames(processes)[colnames(processes)=="resource.x"] <- "resource"
colnames(processes)[colnames(processes)=="resource.y"] <- "mainOutput"

processes <- processes[,c("process",
			 "process.name", 
			 "capacity", 
			 "TRL", 
			 "resource",
			 "unit",
			 "value",
			 "sign",
			 "mainOutput")]
processes <- unique(processes) # TODO: fudge required, due to the duplication of each row when using inner_join (not sure why), so try and fix this.

write.csv(processes, "04_database/shiny-processes.csv")

```

Our database currently contains `r length(unique(processes$process))` unique processes (or `r length(unique(processes$process.name))` when the different capacities are counted). These processes all manage energy, water and waste, encompassing `r length(unique(processes$resource))` resource types, of which 25 are main resources. Processes where energy resources are the main output include fossil fuel- and renewable-based generation technologies (such as power plants and concentrating solar power), as well as methods to produce biofuels and hydrogen. Where water resources are the main output, processes include the delivery and treatment of groundwater and surface water for potable and non-potable purposes. The waste management processes include treatment and energy extraction methods for many waste types including wastewaters, agricultural wastes, municipal solid waste and landfill leachate. Of the `r length(unique(processes$process))` unique processes, `r length(unique(filter(processes, TRL %in% c("c", "r"))$process))` are current, although `r length(unique(filter(processes, TRL=="r")$process))` of these have been identified as having low levels of market penetration (rare). The remaining `r length(unique(filter(processes, !TRL %in% c("c", "r"))$process))` are still in the research and development stage.

### Process comparison

We have argued that a process-oriented perspective on urban metabolism would offer insight on how changing a system's mix of processes could meet demand for energy, water and waste management, but with reduced negative impacts. To appreciate how our database could help realise this, we first consider how processes meet demand for a single resource, using the subset of processes whose main output is methane as an example. Figure \ref{fig:heatmap_CH4} visualises the resource inputs and outputs for six processes to produce 1 kg of methane. At a glance, the proximity of points to the $x=0$ intercept makes evident which production methods have the smallest of the various input and output flows. Coupling these flow values with the impact categories used in life cycle analysis and/or the constraints associated with a specific area would identify the most desirable process choice under certain circumstances. For example, a city which uses the very water-reliant 'Biogas upgrading -- AWR' method may like to consider lower water alternatives. Alternatively, another city may have a ready supply of biomass (perhaps from agricultural waste), which would otherwise go to waste, but could instead feed the gasification process. This latter example is one of the 'medium-term' processes, and therefore would require research and investment if it was to be implemented.

```{r energy-heatmap, fig.height=6, fig.cap="Resource consumption (negative values) and production (positive values) for different processes producing 1 kg of methane. Key: \\textsf{AWR} = alkaline with regeneration, \\textsf{BABIU} = bottom ash upgrading, \\textsf{HPWS} = high pressure water scrubbing; text labels indicate the TRL for each process.\\label{fig:heatmap_CH4}"}

## Extract subset of processes which produce methane as a main output
processes_CH4 <- filter(processes,
			mainOutput=="methane",
			capacity %in% c(0.24, 0.0001319, 0.122))

## Add more reader-friendly process and resource names for the plot
plot_names_processes <- read.csv("04_database/plot_names_processes.csv")
plot_names_resources <- read.csv("04_database/plot_names_resources.csv")

processes_CH4 <- inner_join(processes_CH4, plot_names_processes, by="process")
processes_CH4 <- inner_join(processes_CH4, plot_names_resources, by="resource")

processes_CH4$TRL <- factor(processes_CH4$TRL,
			    levels = c("c", "r", "m", "l")) # More appropriate order for the facets of the plot

## Add more descriptive TRL labels to the database
TRL <- c("c", "r", "m", "l")
TRL_name <- c("current", "rare", "medium-term", "long-term")
TRL_name_lookup <- data.frame(TRL, TRL_name)
processes_CH4 <- inner_join(processes_CH4, TRL_name_lookup, by="TRL")

en_dot <- ggplot(processes_CH4, aes(value, plot_name.y)) +
  theme_bw() +
  theme(axis.text.y = element_text(size=8),
	axis.title.y = element_blank(),
	strip.text = element_text(size=8)) +
  geom_point() +
  xlab("Quantity [kg or MJ]") +
  geom_vline(xintercept=0) +
  facet_wrap(~ plot_name.x, ncol=2) +
  geom_text(aes(x=-23, y=1.5, label=TRL_name), size=3, hjust=0)
en_dot
```

### Synergies in process networks

The methane production example demonstrates the resource transfers which can take place between the energy, water, waste and other sectors. Figure \ref{fig:heatmap_CH4} shows how methane production (an energy resource) can result in energy flows (such as electricity or heat requirements), water flows (such as water inputs) and waste flows (such as vegetable waste inputs and digestate outputs). Inter-sectoral links such as these occur frequently in our database, and are summarised in Figure \ref{fig:sna_interactions}, which reveals numerous processes that link the energy, water, waste and other sectors. These links are important because they mean that collaboration between different firms and service providers can reduce a city's overall resource flows at the top, by using the wastes and by-products of one process as inputs to another (advocated by many including @Meijer2011 and @Leduc2013). For a simple example, consider the material flows in an Indian village reported by @Kestemont2010. A material flow analysis of the village indicates it has both biomass waste outputs and electricity imports. Searching our database, we found 22 processes which have biomass inputs and electricity outputs (including biomass-fuelled CHP plants, straightforward combustion or biomass-to-ethanol conversion processes and in the medium term, fuel-cells). The village could invest in one of these processes, to reduce their expenditure on electricity imports. Some options could even produce other resources, such as methane or heat which can be used elsewhere in the village or sold. However, these benefits must be traded off against the resource input requirements (e.g. for coal or water) and economic decisions (e.g. payback time).

Another way to identify synergies is inspired by @Zhu2014's matchmaking proposal using the E-PRTR dataset introduced earlier. When searching the E-PRTR dataset, we find that methane and CO$_2$ are both common pollutants (produced by 1,623 and 2,153 European industrial facilities respectively, from a total of 32,368 facilities). Our database lists 13 processes which require methane or CO$_2$ inputs, and thus could redirect harmful emissions into useful purposes such as electricity production, algae cultivation and wastewater treatment.


```{r interaction-network}

# Inter-sectoral plot for whole network stored separately, so it's easy to suppress from paper compilation (to save time)
base::source("04_database/resource-network.R")

resource_sectors <- read.csv("04_database/resources-sector.csv")
sectors_lookup <- rbind(c("energy", "blue"),
			c("water", "red"),
			c("waste", "black"),
			c("other", "green"))
sectors_lookup <- data.frame(sectors_lookup)
colnames(sectors_lookup) <- c("sector", "node_vector")
resources_sectors_node <- inner_join(resource_sectors, sectors_lookup)
node_vector <- c(resources_sectors_node$node_vector)

## Network plot (using SNA package)
layout <-  c("fruchtermanreingold", "kamadakawai", "spring", "circle", "eigen", "hall", "mds", "princoord", "target")

#library(sna) # Code to plot figure for saving (and subsequent manual editing)
#gplot(link_matrix,
#      mode=layout[1],
#      gmode="graph",
#      label = resources,
#      label.pos = 1,
#      jitter=FALSE,
##      arrowhead = FALSE,
#      thresh=0,
##      vertex.sides = 0,
#      vertex.col = node_vector,
#      edge.col = rgb(220, 217, 217, maxColorValue=255),
#      displaylabels = TRUE)
```

\begin{figure}
	\centering
	\begin{spacing}{1.0}
	\includegraphics[width=\columnwidth]{04_database/final-network.pdf}
	\end{spacing}
	\caption{The interactions of resources in the database. The proximity of any pair of connected resources is proportional to the number of processes for which one resource of the pair is an input, and the other is an output. Graph is plotted using the R package from \citet{SNA2005}. A colour version of this plot is available in the electronic version of this paper.}
	\label{fig:sna_interactions}
\end{figure}

### Optimisation of networks

Considering the possibilities for inter-sectoral synergies across a whole network of processes is impossible at a glance. When multiple locations in a city are demanding water, heating and electricity as well as generating waste and wastewater, then process-oriented perspectives could adopt optimisation modelling. This technique has been previously used in urban resource planning, including energy systems (e.g. @Keirstead2012), water supply networks (e.g. @Lim2010), and waste management activities (e.g. @Li2006). A tool described in @Ravalde2015 seeks to optimise the management of energy, water and waste simultaneously. With it, a user specifies the time and location of resource demands and waste generation rates, and a mathematical optimisation routine identifies mix of processes that meets a city's demands whilst minimising an objective (such as cost, carbon footprint, water footprint, waste generation etc.). Many model objectives can be functions of the aggregate flows at the top, and thus process-mix optimisation can link the top of the urban system to demands at its bottom.

We demonstrate a simple example based on the hypothetical city from @Wolman1965, using his data to derive bottom-up demands for electricity, heating[^wolman_energy] and water (we ignore motor fuel and food), and generation rates for refuse and wastewater. Making all the technologies in our database available, we optimised the city's choice of processes so as to minimise CO$_2$ emissions at the top. Figure \ref{fig:bar_wolman} shows the city's resource flows at the top before and after optimisation, which reduces emissions to zero. This is possible because the model chooses gas-fuelled CHP to meet energy demands (which requires a slight increase in gas imports, but eliminates coal and oil imports). The model also chooses algae cultivation to absorb any remaining emissions. Sewage flows are also reduced while refuse and water flows remain unchanged. These system changes are unlikely in reality because of constraints on land-use (e.g. for algae production) and cost, nevertheless this example demonstrates the potential benefits of process-oriented optimisation modelling.

[^wolman_energy]: We assume that the fossil fuels (coal, natural gas and oil) meet all electricity and heating demand. We also assume that electricity consumption is 1.12 times the heat consumption (this value is calculated from the mean ratio of electrity-to-heat consumption for all cities in the urban metabolism database from @Kennedy2014).

```{r wolman-flows, fig.height=3.6, fig.cap="Flows at the top of Wolman's hypothetical city before and after optimisation modelling suggests an optimal network for the middle. Negative values represent inputs while positive values represent outputs\\label{fig:bar_wolman}."}

wolman <- read.csv("04_database/wolman-flows.csv")
wolman <- melt(wolman)

plot_wolman <- ggplot(wolman, aes(resource, -value/1e6)) +
  theme_bw() +
#  geom_bar(stat="identity") +
  geom_point() +
  geom_hline(xintercept=0) +
  coord_flip() +
  xlab("Resource (units)") + 
  ylab("Value [units x 10^6]") +
  facet_wrap(~variable, ncol=1)
plot_wolman
```

An optimisation model for a city can explore different aspects to process-oriented planning and policy agendas. For example, it can inform urban planning by considering whether decentralised systems (which use more smaller capacity processes close to the point of resource use or generation) manage resources more efficiently than centralised systems. Modelling could also inform how a system's middle should adapt to changes at the bottom, by altering resource demands to simulate behavioural changes and efficiency measures. Finally, the model can help clarify research and investment efforts, by identifying which processes with medium- and long-term TRLs would most benefit a particular urban area.

### Reflections: interacting with other datasets

The three applications above each depend on having other data available: process comparisons require knowledge of resource inputs and outputs in an existing process; searching for synergies requires knowledge of inputs and outputs to multiple processes; and optimisation modelling needs to know the resource demands and waste generation rates at the bottom of an entire system.

Thus a potential obstacle to using our database is the availability of other data. @Davis2010 argues that within IE, data is often "unavailable, inaccessible, incomplete, incompatible, or unreliable" (p.708), and believes that IE can best overcome these obstacles through increased collaboration. Here, our use of other datasets has been limited to the E-PRTR, but it would be profitable to reach beyond this source and use data more associated with urban metabolism to identify synergies between energy conversion, water treatment and waste management processes in existing cities. Similarly, a dataset of demands for electricity, heating, potable and non-potable water, and waste and wastewater management could be established for domestic, commercial and industrial sectors, for urban areas of different types (such as size, maturity and climate).

It would also be beneficial to have some means of linking databases together. @Davis2010 elucidate why, by recapitulating the arguments of @Bush1945 regarding the most useful ways to store scientific knowledge. Rather than traditional hierarchical storage (e.g alphabetical order or the Dewey Decimal System) which forces all information into a single location, @Bush1945 argued that information should be located according to associative trails, i.e. paths "where facts are connected to other associated facts" [@Davis2010, p.711], as exemplified by Wikipedia's use of hyperlinks.

@Davis2010 take on board Bush's concept, and note that it can be facilitated within IE by using Semantic Wikis (of which Enipedia is an example), which store information in both structured and unstructured formats. Unstructured information records prose (such as the history of a particular process); structured information records machine-readable quantitative information (such as a resource flow value, a capacity or a TRL). Information can contain hyperlinks that point to other relevant pages.

The Semantic Web has attracted some scepticism, largely due to pragmatic concerns. Any machine readable information needs to be encoded according to some standard, which may not be agreed upon by the different parts of a community, and conforming to multiple different standards would incur a high financial and resource cost [@Marshall2003]. However, @Herman2007 has observed that the Semantic Web has been successful within specialised communities (such as financial services and health care); thus there is promise for it within IE. Furthermore, to minimise the efforts required by contributors, machine learning has the potential to automate the searching of information to discover structure within it, and then to formally encode this structure with the appropriate Semantic Web code [@Tresp2008].

@Davis2010 illustrate how a Semantic Wiki could assist a hypothetical life cycle analysis researcher. Inspired by this example, we illustrate how urban metabolism research might benefit similarly. Consider how the 'Biogas upgrading -- AWR' processes information (Figure \ref{fig:heatmap_CH4}) might form part of a Semantic Wiki.  Its wiki page would use unstructured data to describe how it works (perhaps including details on the process chemistry), and use structured data to record the resource inputs and outputs. The `Water -- non-potable` flow would be a hyperlink which links to a page about non-potable water. This page could then link to a page which lists the all processes which produce `Water -- non-potable` (both from our database, and other sources), which could then link to a page which lists all the cities which currently use such processes. Thus a researcher can travel through several datasets including our own, to find possible applications of technologies in our database.

## Summary and further work

Both urbanisation trends and problems associated with urban resource consumption make urban metabolism an important field of study. Top-down and bottom-up approaches to urban metabolism offer important insights on a city's aggregate resource consumption and waste generation, and how this depends on the consumption and waste of homes, business, industry and transport. However, to date, a process-oriented view of urban metabolism has received little attention. Thus we built a database comprising 202 unique processes, which manage energy, water and waste, both currently operational and those still in research and development.

We have shown that the main contribution of process-oriented approaches is to inform decision makers how best to plan and invest in processes which improve a city's metabolism. This can be to choose the best of several alternatives to manage a single resource (in the case of a single utility service provider); to choose processes which take advantage of potential synergies in an area (in the case of a cooperative of organisations), or to optimise a city's whole network of processes (in the case of a local government or researcher). The information provided to decision makers will improve as the research community takes advantage of database's open data format to continually improve it, in the following ways:

1. Adding new processes
2. Broadening the type of information each process records, to include process costs (for capital, operation and maintenance), and other pollutants (such as SO$_2$), and other data.
3. Extending the database beyond energy, water and waste management processes, to consider other processes which affect the metabolism of an area, such as steel and concrete manufacturing.

Another area of future work is to support not just the evolution of the database, but of the whole ecosystem in which it lives. This is because we have seen that process-oriented approaches are not isolated from other  approaches, but rather are most useful when using information from them, such as demand data from the bottom. The urban metabolism data ecosystem could flourish in the Semantic Web, which can link together structured and unstructured information in a vast network, through which researchers can navigate, finding innovative and as yet unknown applications for the data. Our database is publicly available under an open-source licence and so we hope that it will become part of an evolving ecosystem of linked urban metabolism data.

In summary, our database can facilitate process-oriented approaches to urban metabolism, especially if it becomes part of an information ecosystem in which data from the top, middle and bottom of urban system are considered together in more holistic urban metabolism assessments. Ultimately, this will help cities meet their inhabitants' needs more sustainability, in response to the challenges presented by urbanisation and climate change.
