# Case study data

[TODO: to update from webpage version]

This webpage gives further detail on the data used in the case study outlined in @Ravalde2016b. Specifically we:

- present summaries of the data
- provide links to the data (in CSV and YAML formats)
- note data sources, along with assumptions and calculations used to convert the data into appropriates values and units
- [TODO: if time] Show how it can be converted to be used in GAMS modelling (using the R programming language)

All of the above can be found from this webpage. Furthermore, since the source of this page was written in [Rmarkdown](http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html) format before its conversion to HTML, by downloading the [supInf.Rmd](supInf.Rmd) source of this page, you can get the R-code used on this page to manipulate the data, for your own interest.

The next section outlines the data in the order in which we presented it in @Ravalde2016b in the subsection titled 'Input data'.

## A note on sources

Some of our data and assumptions come from seven documents (Word documents and PowerPoint presentations) provided by the Shann Gu company and their collaborators. This documentation is not publicly available. We will list these references using a square-bracketed number between 1 and 7, e.g. [3].

Other sources will be referenced by links.

## Input data

### Site layout

Any size used in the model needs to be defined as a number of zones. For the Shann Gu model we use the thirteen zones listed on slide 20 of [6]. Each zone belongs to a region. 

Region, Zone
south, factory
south, factory_assembly
production, northDistrict
production, multifunction
production, factory_blades
production, tech
production, quality
production, dormitories
residential, residential
residential, hospital
residential, school
energyIsland, energyIsland
NA, reserved

Next we define which of the above zones can be directly connected by transport links. This is done from a visual inspection to identify neighbouring zones from slide 20 of [6]. At the same time, we can define the distance between such neighbouring zones using a origin-destination matrix [zones.csv](../modelling/data/zone-matrix.csv). Any origin-destination pair which is are not neighbours are denoted with `NA` in its cell (e.g. `factory-northDistrict`). The other cells contain the distance between zones, in kilometers (e.g. `factory-factory_assembly` has a distance of [TODO: number]). This matrix is lower-triangular, so there is no need to define a distance in both directions (we assume these are equal).

The distances have been assumed as being approximately [TODO: number] km, based on the overall site being 13 km^2.

[TODO: add the code used to define the distances]

### Times

Our case study's time steps (along with their lengths in seconds) are defined in the [times.csv](../modelling/data/times.csv) file.

We have assumed a 360 day year, with the winter, summer and shoulder seasons taking lengths of 90 days, 90 days, and 180 days, respectively. So for example, the length of the `shoulder` season is calculated as $180~\mbox{days} \times 24~\mbox{hours} \times 3,600~\mbox{seconds} = 1.5552 \times 10^7~\mbox{seconds}$.

### Resource demands

Summary of adjustments:

- energy
	- generate shoulder heating
	- distribute energy demands among zones (regions --> functions)
	- attribute energy demands to cities
- other
	- Calculate water (potable), water (reclaimed), wastewater, waste (nonorganic), waste (organic) etc.
	- distribute among zones
	- attribute other demands to cities

SPC [demand] gives the site's energy demands, specifically it gives demands for electricty (year round), heating (winter) and cooling (summer). These demands are defined for each zone, so no adjustments need be applied in this regard. However, we do assume that heating is required in the shoulder season, and assume it to be half the value

The following demands zones are given in the Shann Gu documentation:

- Slide 20 of [6]: `heat` (winter), `cool` (summer) and `elec` (year-round).
- Slide 26 of [5]: `waterPotable`, and `water_reclaimed` as well as rates of `wastewater` generation
- Slide 28 of [5]: rates of `waste` generation (both organic and non-organic)
 
Not that waste and wastewater generation rates are defined as negative demands, for use in our model.

The waste generation figures from the Shann Gu Documentation are not provided in the same units (total solid waste = 8m3/day, organic waste = 360,000 kg/year), so we perform the following calculations to get separate values for `waste` and `waste_organic`:

Assume solid waste has a density $\rho=217.5~\mbox{kg/m3}$[^waste-density], thus:

[^waste-density]: Table 2.2 from [Solid Waste Management by Chandrappa and Das](http://www.springer.com/cda/content/document/cda_downloaddocument/9783642286803-c2.pdf?SGWID=0-0-45-1336211-p174308664) gives a density range of 87--348 kg/m3. We take the median value of 217.5 kg/m3.

$$
\begin{align}
\mbox{Total solid waste} &= 8~\mbox{m3/day} \\
 		  &= 1,740~\mbox{kg/day} \\
 		  &= 635,100~\mbox{kg/year}
\end{align}
$$

Subtracting the organic waste fraction: 

$$
\begin{align}
\mbox{Non-organic waste} &= 635,100 - 360,000~\mbox{tons/year} \\
 		  &= 275,100~\mbox{kg/year}
\end{align}
$$

Having identified the site's resource management demands, we next make assumptions and calculations to appropriately redistribute some demands between multiple zones, before converting all demands to a consistent set of units (MW for energy, kg/s otherwise).

In [resource-demands.csv](resource-demands.csv), we detail both the original data taken from the Shann Gu documentation, and any subsequent assumptions and calculations applied to that data. We now briefly outline this file's columns (denoted by **bold** text):

- **zone.region** and **zone** are those regions and zones referred to in the Site Layout section above. 
- The documented demand for a resource in any given season is denoted using the **resource**, **season**, and **value.doc** columns, with the units and reference given in **units.doc** and **reference**.
- When the Shann Gu documentation gives little detail about the spatial disaggregation of the demand (i.e. demand is only attributed to a region, rather than a zone), we make an assumption about how to divide this demand up between zones. Similarly, when the documentation has not provided a demand value for a particular season, we make assumptions (e.g. we have assumed that the shoulder heat demand is half the winter heat demand). This information is recorded in **assump.redist**, with the new values listed in **value.redist** (and units given in **units.redist**)
- Finally, **notes.conv** notes any necessary changes in units, such that **value.conv** and **units.conv** contain resource demands in a consistent set of units.
<!--
- More notes and calculations (for authors' use) in data-notes.ods.

Example of how to convert documented electricity demand (kWh/year) to units required by the model (MW):
$$
\begin{align}
111700000 \times 3.6/(365 \times 24 \times 3600) &= 12.75 \\
6e7 \times 3.6/(365 \times 24 \times 3600) &= 6.85
\end{align}
-->

In summary, the reasoning for every demand value which appears in our case study can be justified by the chain of sources, assumptions, and calculations embedded in our CSV file.

The data is recorded in CSV format in [resource-demands.csv](resource-demands.csv). Below are all rows from this dataset in which the **value.conv** column contains a non-zero value (i.e. all the resource demands to be in our case study), along with the R required to extract these rows.

```{r data-head, echo="TRUE"}
demands <- read.csv("resource-demands.csv")

library("dplyr")
filter(demands, value.conv!=0)

## Arranged according to zone
#demands_nonzero <- demands %>% arrange(zone.number) %>% select(zone, resource, season, value.conv, units.conv) %>% filter(value.conv!=0)
#
## Save for use in GAMS model
#write.csv(demands_nonzero, "../modelling/data/demands-nonzero.csv", row.names=FALSE)
```

### Resource options

As well as the resource management demands specified above, our case study considers other resources which can play a part in the conversion chain of processes which meet demand (e.g. coal can be used in a power plant which also requires cooling water, and emits carbon dioxide).

These resources are defined using the YAML format, by which which define a resource's name (e.g. `coal`), and also define quality attributes $X_{rq}$ (e.g. `mass`), for example:

```yaml
resource: coal
quality:
        mass: 1
notes:
```

The YAML files for all [TODO: number] resources for our model are available [here](../modelling/data/resources-yaml). [TODO: link to GitHub, or convert to zip?]

For each resource, we also need to define parameters including its seasonal import and export limits, the number of zones allowed to import and export it, and its cost. We tabulate all this in CSV format: [resource-params.csv](../modelling/data/resource-params.csv). This data also includes sources, assumptions and calculation steps used to estimate resource costs. The top few rows of this data set are shown below.


```{r data-params-head, echo="TRUE"}
params <- read.csv("../modelling/data/resource-params.csv")
head(params)
```

### Process options

We also use YAML files to define the characteristics of the processes which convert and transport resources, as outlined below.

#### Conversion processes

For each conversion process, a YAML file defines the relative quantities of input and output `resources`, and the maximum capacity (or `rate`) at which a process can operate. For example, a coal-fuelled power plant is defined thus:

```yaml
process: elecplant-coal-l
resources:
     coal: -0.1012658
     elec: 1
     heat: 1.56962
     water: -0.8670886
     CO2: 0.2341772
     ash: 0.01012658
rate:
        wint: 5400
        sum: 5400
        shoulder: 5400
notes:
```

The processes main output (`elec` in this case) is always set to a value of `1`, such that the `rate` multiplied by any resource flow will give the maximum input or output of any resource.

The YAML files for all the processes used in our case study are available [here](../modelling/data/processes-yaml). [TODO: link to GitHub, or convert to zip?]. This data comes from two sources:

- a few processes have been defined using information from the Shann Gu literature
- the remaining majority of process definitions are adapted from the publicly available [database of urban metabolism processes](https://github.com/tomravalde/urban-metabolism-process-database) which is introduced in @Ravalde2016.

We also define each process's cost in [process-costs.csv](../modelling/data/process-costs.csv). As discussed in the paper[^process-costs], various factors make estimating process costs tricky. However, to run the model, it is helpful to have plausible process costs.

[^process-costs]: Considerable effort is required to research the costs of each process, especially given that some technologies are still under research -- economic forecasting is beyond our scope. Even apart from these unknowns, there are other complicating factors, including variations in purchasing power parities and exchange rates, as well as the different environmental constraints/conditions into which a process is built.

Our approach is to build a database of general processes ($GP$) for which we can have a unit cost (USD/MW or USD/kg/s) for each process. We can then assign each model process ($MP$) with a $GP$ (indicated by the [TODO: relevant column] column in process-costs.csv). For processes which do not yet exist [TODO: e.g. <example>], we use a GP which produces/consumes a similar main resource. Once each MP has an associated GP Unit Cost, we can estimate the an MP Total Cost (USD) by calculating:

$$
\begin{align}
\mbox{MP Total Cost} = \mbox{GP Unit Cost} \times \mbox{MP Flow Rate}
\end{align}
$$

So that our modelling accounts for economies of scale, each GP has both lowerbound and upperbound unit costs, with the former being applied to larger instances of a process, and the latter applied to a smaller instance. To estimate $\mbox{GP Unit Cost}$, we research the literature, and apply appropriate assumption, conversions and calculations, all of which can be traced in [process-cost-calcs-kgyr.ods](../modelling/data/process-cost-calcs-kgyr.ods). The terms in **bold** refer to column headers in the spreadsheet.

1. From our list of MPs, we assemble a list of suitable GPs (which can apply to more than one MP)
2. For each GP, we list a lowerbound and upperbound versions, appending `_l` and `_u` to the process names, respectively (e.g. `CHP-biogas_l` and `CHP-biogas_u`). These GPs are listed in the **process-type** column.
3. We find costs for each process type (indicating this using **ref.cost** and **ref.currency**, with the **reference** column providing a link to the source)
4. If necessary we convert this value to USD in the **ref.USD** column.
5. The referenced cost will correspond to a particular resource and its production (or consumption) rate. For example, a CHP cost may be per kW of electricity produced. The resource, its quantity, and its units, are recorded in **ref.flow**, **ref.flow.value** and **ref.flow.units**, respectively.
6. We can then calculate a **ref.flow.mult.factor** which is the number by which the **ref.flow.value** should be multiplied, in order to convert the flow into a consistent set of units (MW for energy resources, kg/s otherwise). Any relevant notes (e.g. the density of a material) are recorded in **mult.factor.notes**.
7. Finally we can calculate a **ref.unit.cost** each process-type (USD/MW or USD/kg/s) by calculating **ref.USD**/(**ref.flow.value** $\times$ **ref.flow.mult.factor**).

#### Transport processes

For each transport process, a YAML file defines the resource carried between an origin and destination using `inputs` and `outputs` respectively. In the example of coal transportation below, now coal is lost between the origin and the destination. The `rate` mapping defines the maximum transport rate (MW or kg/s). The YAML file also defines the quantity of any resource consumed/produced in that activity, in this case gasoline consumption and CO2 emissions. We split these values equally between the origin and destination zones.

```yaml
process: road_coal
inputs:
        coal: -1
        gasoline: 0
        CO2: 0
outputs:
        coal: 1
        gasoline: 0
        CO2: 0
dist_source:
        coal: 0
        gasoline: -0.18
        CO2: 0.3
dist_dest:
        coal: 0
        gasoline: 0
        CO2: 0.3
rate: 100
notes:
```

The remaining transport processes can be found in [transport-yaml](../modelling/data/transport-yaml). We also provide [calculation details](transp-process-calcs.html) for electricity, pipe flow (water), pipe flow (gas), and vehicle transport).

The transport costs are defined per kilometer in [transport.csv](../modelling/data/transport.csv), which also gives references.

## Model scenarios

- All the above data can be subsetted according to appropriate scenarios
- YAML configuration files define which subset of data to be used
- Our R script then turns this into a model instance

### List the YAML files for each scenario in our paper

## Summary

Summary text.
